---
title: "BayeScan input files"
author: "Grace Smith-Vidaurre"
date: "September 14, 2020"
output: html_document
---

```{r setup, eval = TRUE, echo = FALSE}

knitr::opts_knit$set(root.dir = "/home/owner/Desktop/GitHub_repos/origins-selection")

```

Purpose: Subset the Structure file of the HWE and missing data filtered merged SNPs to make versions representing all pairwise population comparisons to be executed in BayeScan. The files generated in this script will be in Structure format, and will be converted to BayeScan formt using a script to execute PGDSpider.

Note that I did not filter out positive controls within or across libraries here. Need to do some validations of positive controls in the exploratory visualizations, and filter these out if possible, then remake these BayeScan input files.
```{r echo = TRUE, eval = TRUE, message = FALSE}

rm(list = ls())

X <- c("tidyverse", "pbapply", "data.table", "adegenet", "openxlsx")
invisible(lapply(X, library, character.only = TRUE))

# Path to the metadata spreadsheet that will be updated
xls_path <- "/home/owner/Desktop/MANUSCRIPTS/Origins_Selection/DATA/Metadata_Barcodes"

# Path to Stacks output, including the HWE filtered SNPs in Structure format 
res_path <- "/media/owner/MYIOPSITTA/R/Origins_Selection/stacks"

gpath <- "/home/owner/Desktop/MANUSCRIPTS/Origins_Selection/GRAPHICS"

seed <- 401

```

Read in metadata.
```{r echo = TRUE, eval = TRUE}

meta_dats <- read.xlsx(file.path(xls_path, "Mymon_RADlibraries_2015_2019_CombinedMetadata.xlsx"))
glimpse(meta_dats)

```

Read in the HWE and missing data filtered merged SNPs.
```{r echo = TRUE, eval = TRUE}

struc <- read.table(file.path(file.path(res_path, "merged"), "merged_HWE_missingData_filters.str"), skip = 1)
# glimpse(struc)
dim(struc) # 432 rows, 563 columns
head(struc[, 1]) # Sample names contained in the first column
head(struc[, 2]) # Population ID (single population) contained in second column
names(struc)

# Get the column headers

# Read the first line of metadata, these are the column names
col_nms <- readLines(file.path(file.path(res_path, "merged"), "merged_HWE_missingData_filters.str"), n = 1)
col_nms

# Split out the column headers into a vector, one element per header
mrkrs <- strsplit(strsplit(col_nms, split = "\t")[[1]][3], split = "[[:space:]]")[[1]]
head(mrkrs)
length(mrkrs)

# Get rid of the space that precedes the first locus column header
mrkrs <- mrkrs[-which(mrkrs == "")]
head(mrkrs)
length(mrkrs)

# Add the individual and population column headers that precede the locus column names
col_nms <- c("indiv", "pop", mrkrs)
head(col_nms)
length(col_nms)

# Add back the updated column names
# Looks good
names(struc) <- col_nms
glimpse(struc[, 1:10])

```

These are the general population contrasts I'm considering. Some of these will entail multiple files/comparisons (e.g. 14-18):

1) All invasive sites versus all native sites
2) All invasive sites versus all Uruguay sites
3) All invasive sites versus all Argentina sites
4) All invasive sites versus all northern Argentina sites
5) All invasive sites versus all southern Argentina sites
6) All U.S. sites versus all Uruguay sites
7) All U.S. sites versus all Argentina sites
8) All U.S. sites versus all northern Argentina sites
9) All U.S. sites versus all southern Argentina sites
10) All Spain sites versus all Uruguay sites
11) All Spain sites versus all Argentina sites
12) All Spain sites versus all northern Argentina sites
13) All Spain sites versus all southern Argentina sites
14) All United States versus all Spain sites
15) Each invasive site versus all native sites (9 comparisons)
16) Each invasive site versus all Uruguay sites (9 comparisons)
17) Each invasive site versus all Argentina sites (9 comparisons)
18) Each invasive site versus all northern Argentina sites (9 comparisons)
19) Each invasive site versus all southern native sites (9 comparisons)
20) Randomizations


1) All invasive sites versus all native sites. Change the population column encode invasive versus native.
```{r echo = TRUE, eval = TRUE}

struc_tmp <- struc %>%
  inner_join(
    meta_dats %>%
      dplyr::select(Sample_Name, Population),
    by = c("indiv" = "Sample_Name")
  ) %>%
  dplyr::mutate(
    pop = factor(Population, levels = c("Native", "Invasive")),
    pop = as.numeric(pop)
  ) %>%
  dplyr::select(-c(Population))

# Looks good
glimpse(struc_tmp[, 1:10])
struc_tmp %>%
  slice(1:50) %>%
  dplyr::select(indiv, pop) %>%
  View()

# Write a new file
file_nm <- "BS_01_AllInvasiveVersusNative.str"
file.remove(file.path(file.path(res_path, "merged"), file_nm))

header <- c("\t","\t", paste(paste(mrkrs, collapse = " "), "\n", sep = ""))

cat(header, file = file.path(file.path(res_path, "merged"), file_nm), append = TRUE)

invisible(pbsapply(1:nrow(struc_tmp), function(i){
  
  cat(c(struc_tmp$indiv[i], "\t", struc_tmp$pop[i], "\t", paste(paste(struc_tmp[i, -grep("^indiv$|^pop$", names(struc_tmp))], collapse = " "), "\n", sep = "")), file = file.path(file.path(res_path, "merged"), file_nm), append = TRUE)
  
}))

# Opened the file in Vim to doublecheck structure, looks good
# Use :set list in Vim to see invisible characters (tabs)

```

2) All invasive sites versus all Uruguay sites.
```{r echo = TRUE, eval = TRUE}

struc_tmp <- struc %>%
  inner_join(
    meta_dats %>%
      dplyr::select(Sample_Name, Population, Country),
    by = c("indiv" = "Sample_Name")
  ) %>%
  dplyr::mutate(
    pop = factor(Population, levels = c("Native", "Invasive")),
    pop = as.numeric(pop)
  ) %>%
  filter(Country %in% c("United States", "Spain", "Uruguay")) %>%
  dplyr::select(-c(Population, Country))

dim(struc_tmp) # 173 individuals

# Looks good
glimpse(struc_tmp[, 1:10])
struc_tmp %>%
  dplyr::select(indiv, pop) %>%
  View()

# Write a new file
file_nm <- "BS_02_AllInvasiveVersusUruguay.str"
file.remove(file.path(file.path(res_path, "merged"), file_nm))

header <- c("\t","\t", paste(paste(mrkrs, collapse = " "), "\n", sep = ""))

cat(header, file = file.path(file.path(res_path, "merged"), file_nm), append = TRUE)

invisible(pbsapply(1:nrow(struc_tmp), function(i){
  
  cat(c(struc_tmp$indiv[i], "\t", struc_tmp$pop[i], "\t", paste(paste(struc_tmp[i, -grep("^indiv$|^pop$", names(struc_tmp))], collapse = " "), "\n", sep = "")), file = file.path(file.path(res_path, "merged"), file_nm), append = TRUE)
  
}))

# Opened the file in Vim to doublecheck structure, looks good

```

3) All invasive sites versus all Argentina sites
```{r echo = TRUE, eval = TRUE}

struc_tmp <- struc %>%
  inner_join(
    meta_dats %>%
      dplyr::select(Sample_Name, Population, Country),
    by = c("indiv" = "Sample_Name")
  ) %>%
  dplyr::mutate(
    pop = factor(Population, levels = c("Native", "Invasive")),
    pop = as.numeric(pop)
  ) %>%
  filter(Country %in% c("United States", "Spain", "Argentina")) %>%
  dplyr::select(-c(Population, Country))

dim(struc_tmp) 
nrow(struc_tmp)/2 # 169 individuals

# Looks good
glimpse(struc_tmp[, 1:10])
struc_tmp %>%
  dplyr::select(indiv, pop) %>%
  View()

# Write a new file
file_nm <- "BS_03_AllInvasiveVersusArgentina.str"
file.remove(file.path(file.path(res_path, "merged"), file_nm))

header <- c("\t","\t", paste(paste(mrkrs, collapse = " "), "\n", sep = ""))

cat(header, file = file.path(file.path(res_path, "merged"), file_nm), append = TRUE)

invisible(pbsapply(1:nrow(struc_tmp), function(i){
  
  cat(c(struc_tmp$indiv[i], "\t", struc_tmp$pop[i], "\t", paste(paste(struc_tmp[i, -grep("^indiv$|^pop$", names(struc_tmp))], collapse = " "), "\n", sep = "")), file = file.path(file.path(res_path, "merged"), file_nm), append = TRUE)
  
}))

# Opened the file in Vim to doublecheck structure, looks good

```

4) All invasive sites versus all northern Argentina sites
```{r echo = TRUE, eval = TRUE}

struc_tmp <- struc %>%
  inner_join(
    meta_dats %>%
      dplyr::select(Sample_Name, Population, Country, Site_Code),
    by = c("indiv" = "Sample_Name")
  ) %>%
  dplyr::mutate(
    pop = factor(Population, levels = c("Native", "Invasive")),
    pop = as.numeric(pop)
  ) %>%
  filter(Country %in% c("United States", "Spain") | Country == "Argentina" & !Site_Code %in% c("ALGA", "BAIR", "LURO")) %>%
  dplyr::select(-c(Population, Country, Site_Code))

dim(struc_tmp) 
nrow(struc_tmp)/2 # 144 individuals

# Looks good
glimpse(struc_tmp[, 1:10])
struc_tmp %>%
  dplyr::select(indiv, pop) %>%
  View()

# Write a new file
file_nm <- "BS_04_AllInvasiveVersusNorthArgentina.str"
file.remove(file.path(file.path(res_path, "merged"), file_nm))

header <- c("\t","\t", paste(paste(mrkrs, collapse = " "), "\n", sep = ""))

cat(header, file = file.path(file.path(res_path, "merged"), file_nm), append = TRUE)

invisible(pbsapply(1:nrow(struc_tmp), function(i){
  
  cat(c(struc_tmp$indiv[i], "\t", struc_tmp$pop[i], "\t", paste(paste(struc_tmp[i, -grep("^indiv$|^pop$", names(struc_tmp))], collapse = " "), "\n", sep = "")), file = file.path(file.path(res_path, "merged"), file_nm), append = TRUE)
  
}))

# Opened the file in Vim to doublecheck structure, looks good

```

5) All invasive sites versus all southern Argentina sites
```{r echo = TRUE, eval = TRUE}

struc_tmp <- struc %>%
  inner_join(
    meta_dats %>%
      dplyr::select(Sample_Name, Population, Country, Site_Code),
    by = c("indiv" = "Sample_Name")
  ) %>%
  dplyr::mutate(
    pop = factor(Population, levels = c("Native", "Invasive")),
    pop = as.numeric(pop)
  ) %>%
  filter(Country %in% c("United States", "Spain") | Country == "Argentina" & Site_Code %in% c("ALGA", "BAIR", "LURO")) %>%
  dplyr::select(-c(Population, Country, Site_Code))

dim(struc_tmp) 
nrow(struc_tmp)/2 # 152 individuals

# Looks good
glimpse(struc_tmp[, 1:10])
struc_tmp %>%
  dplyr::select(indiv, pop) %>%
  View()

# Write a new file
file_nm <- "BS_05_AllInvasiveVersusSouthArgentina.str"
file.remove(file.path(file.path(res_path, "merged"), file_nm))

header <- c("\t","\t", paste(paste(mrkrs, collapse = " "), "\n", sep = ""))

cat(header, file = file.path(file.path(res_path, "merged"), file_nm), append = TRUE)

invisible(pbsapply(1:nrow(struc_tmp), function(i){
  
  cat(c(struc_tmp$indiv[i], "\t", struc_tmp$pop[i], "\t", paste(paste(struc_tmp[i, -grep("^indiv$|^pop$", names(struc_tmp))], collapse = " "), "\n", sep = "")), file = file.path(file.path(res_path, "merged"), file_nm), append = TRUE)
  
}))

# Opened the file in Vim to doublecheck structure, looks good

```

6) All U.S. sites versus all Uruguay sites
```{r echo = TRUE, eval = TRUE}

struc_tmp <- struc %>%
  inner_join(
    meta_dats %>%
      dplyr::select(Sample_Name, Population, Country),
    by = c("indiv" = "Sample_Name")
  ) %>%
  dplyr::mutate(
    pop = factor(Population, levels = c("Native", "Invasive")),
    pop = as.numeric(pop)
  ) %>%
  filter(Country %in% c("United States", "Uruguay")) %>%
  dplyr::select(-c(Population, Country))

dim(struc_tmp) 
nrow(struc_tmp)/2 # 133 individuals

# Looks good
glimpse(struc_tmp[, 1:10])
struc_tmp %>%
  dplyr::select(indiv, pop) %>%
  View()

# Write a new file
file_nm <- "BS_06_UnitedStatesVersusUruguay.str"
file.remove(file.path(file.path(res_path, "merged"), file_nm))

header <- c("\t","\t", paste(paste(mrkrs, collapse = " "), "\n", sep = ""))

cat(header, file = file.path(file.path(res_path, "merged"), file_nm), append = TRUE)

invisible(pbsapply(1:nrow(struc_tmp), function(i){
  
  cat(c(struc_tmp$indiv[i], "\t", struc_tmp$pop[i], "\t", paste(paste(struc_tmp[i, -grep("^indiv$|^pop$", names(struc_tmp))], collapse = " "), "\n", sep = "")), file = file.path(file.path(res_path, "merged"), file_nm), append = TRUE)
  
}))

# Opened the file in Vim to doublecheck structure, looks good

```

7) All U.S. sites versus all Argentina sites
```{r echo = TRUE, eval = TRUE}

struc_tmp <- struc %>%
  inner_join(
    meta_dats %>%
      dplyr::select(Sample_Name, Population, Country),
    by = c("indiv" = "Sample_Name")
  ) %>%
  dplyr::mutate(
    pop = factor(Population, levels = c("Native", "Invasive")),
    pop = as.numeric(pop)
  ) %>%
  filter(Country %in% c("United States", "Argentina")) %>%
  dplyr::select(-c(Population, Country))

dim(struc_tmp) 
nrow(struc_tmp)/2 # 128 individuals

# Looks good
glimpse(struc_tmp[, 1:10])
struc_tmp %>%
  dplyr::select(indiv, pop) %>%
  View()

# Write a new file
file_nm <- "BS_07_UnitedStatesVersusArgentina.str"
file.remove(file.path(file.path(res_path, "merged"), file_nm))

header <- c("\t","\t", paste(paste(mrkrs, collapse = " "), "\n", sep = ""))

cat(header, file = file.path(file.path(res_path, "merged"), file_nm), append = TRUE)

invisible(pbsapply(1:nrow(struc_tmp), function(i){
  
  cat(c(struc_tmp$indiv[i], "\t", struc_tmp$pop[i], "\t", paste(paste(struc_tmp[i, -grep("^indiv$|^pop$", names(struc_tmp))], collapse = " "), "\n", sep = "")), file = file.path(file.path(res_path, "merged"), file_nm), append = TRUE)
  
}))

# Opened the file in Vim to doublecheck structure, looks good

```

8) All U.S. sites versus all northern Argentina sites
```{r echo = TRUE, eval = TRUE}

struc_tmp <- struc %>%
  inner_join(
    meta_dats %>%
      dplyr::select(Sample_Name, Population, Country, Site_Code),
    by = c("indiv" = "Sample_Name")
  ) %>%
  dplyr::mutate(
    pop = factor(Population, levels = c("Native", "Invasive")),
    pop = as.numeric(pop)
  ) %>%
  filter(Country == "United States" | Country == "Argentina" & !Site_Code %in% c("ALGA", "BAIR", "LURO")) %>%
  dplyr::select(-c(Population, Country, Site_Code))

dim(struc_tmp) 
nrow(struc_tmp)/2 # 103 individuals

# Looks good
glimpse(struc_tmp[, 1:10])
struc_tmp %>%
  dplyr::select(indiv, pop) %>%
  View()

# Write a new file
file_nm <- "BS_08_UnitedStatesVersusNorthArgentina.str"
file.remove(file.path(file.path(res_path, "merged"), file_nm))

header <- c("\t","\t", paste(paste(mrkrs, collapse = " "), "\n", sep = ""))

cat(header, file = file.path(file.path(res_path, "merged"), file_nm), append = TRUE)

invisible(pbsapply(1:nrow(struc_tmp), function(i){
  
  cat(c(struc_tmp$indiv[i], "\t", struc_tmp$pop[i], "\t", paste(paste(struc_tmp[i, -grep("^indiv$|^pop$", names(struc_tmp))], collapse = " "), "\n", sep = "")), file = file.path(file.path(res_path, "merged"), file_nm), append = TRUE)
  
}))

# Opened the file in Vim to doublecheck structure, looks good

```

9) All U.S. sites versus all southern Argentina sites
```{r echo = TRUE, eval = TRUE}

struc_tmp <- struc %>%
  inner_join(
    meta_dats %>%
      dplyr::select(Sample_Name, Population, Country, Site_Code),
    by = c("indiv" = "Sample_Name")
  ) %>%
  dplyr::mutate(
    pop = factor(Population, levels = c("Native", "Invasive")),
    pop = as.numeric(pop)
  ) %>%
  filter(Country == "United States" | Country == "Argentina" & Site_Code %in% c("ALGA", "BAIR", "LURO")) %>%
  dplyr::select(-c(Population, Country, Site_Code))

dim(struc_tmp) 
nrow(struc_tmp)/2 # 111 individuals

# Looks good
glimpse(struc_tmp[, 1:10])
struc_tmp %>%
  dplyr::select(indiv, pop) %>%
  View()

# Write a new file
file_nm <- "BS_09_UnitedStatesVersusSouthArgentina.str"
file.remove(file.path(file.path(res_path, "merged"), file_nm))

header <- c("\t","\t", paste(paste(mrkrs, collapse = " "), "\n", sep = ""))

cat(header, file = file.path(file.path(res_path, "merged"), file_nm), append = TRUE)

invisible(pbsapply(1:nrow(struc_tmp), function(i){
  
  cat(c(struc_tmp$indiv[i], "\t", struc_tmp$pop[i], "\t", paste(paste(struc_tmp[i, -grep("^indiv$|^pop$", names(struc_tmp))], collapse = " "), "\n", sep = "")), file = file.path(file.path(res_path, "merged"), file_nm), append = TRUE)
  
}))

# Opened the file in Vim to doublecheck structure, looks good

```

10) All Spain sites versus all Uruguay sites
```{r echo = TRUE, eval = TRUE}

struc_tmp <- struc %>%
  inner_join(
    meta_dats %>%
      dplyr::select(Sample_Name, Population, Country),
    by = c("indiv" = "Sample_Name")
  ) %>%
  dplyr::mutate(
    pop = factor(Population, levels = c("Native", "Invasive")),
    pop = as.numeric(pop)
  ) %>%
  filter(Country %in% c("Spain", "Uruguay")) %>%
  dplyr::select(-c(Population, Country))

dim(struc_tmp) 
nrow(struc_tmp)/2 # 88 individuals

# Looks good
glimpse(struc_tmp[, 1:10])
struc_tmp %>%
  dplyr::select(indiv, pop) %>%
  View()

# Write a new file
file_nm <- "BS_10_SpainVersusUruguay.str"
file.remove(file.path(file.path(res_path, "merged"), file_nm))

header <- c("\t","\t", paste(paste(mrkrs, collapse = " "), "\n", sep = ""))

cat(header, file = file.path(file.path(res_path, "merged"), file_nm), append = TRUE)

invisible(pbsapply(1:nrow(struc_tmp), function(i){
  
  cat(c(struc_tmp$indiv[i], "\t", struc_tmp$pop[i], "\t", paste(paste(struc_tmp[i, -grep("^indiv$|^pop$", names(struc_tmp))], collapse = " "), "\n", sep = "")), file = file.path(file.path(res_path, "merged"), file_nm), append = TRUE)
  
}))

# Opened the file in Vim to doublecheck structure, looks good

```

11) All Spain sites versus all Argentina sites
```{r echo = TRUE, eval = TRUE}

struc_tmp <- struc %>%
  inner_join(
    meta_dats %>%
      dplyr::select(Sample_Name, Population, Country),
    by = c("indiv" = "Sample_Name")
  ) %>%
  dplyr::mutate(
    pop = factor(Population, levels = c("Native", "Invasive")),
    pop = as.numeric(pop)
  ) %>%
  filter(Country %in% c("Spain", "Argentina")) %>%
  dplyr::select(-c(Population, Country))

dim(struc_tmp) 
nrow(struc_tmp)/2 # 83 individuals

# Looks good
glimpse(struc_tmp[, 1:10])
struc_tmp %>%
  dplyr::select(indiv, pop) %>%
  View()

# Write a new file
file_nm <- "BS_11_SpainVersusArgentina.str"
file.remove(file.path(file.path(res_path, "merged"), file_nm))

header <- c("\t","\t", paste(paste(mrkrs, collapse = " "), "\n", sep = ""))

cat(header, file = file.path(file.path(res_path, "merged"), file_nm), append = TRUE)

invisible(pbsapply(1:nrow(struc_tmp), function(i){
  
  cat(c(struc_tmp$indiv[i], "\t", struc_tmp$pop[i], "\t", paste(paste(struc_tmp[i, -grep("^indiv$|^pop$", names(struc_tmp))], collapse = " "), "\n", sep = "")), file = file.path(file.path(res_path, "merged"), file_nm), append = TRUE)
  
}))

# Opened the file in Vim to doublecheck structure, looks good

```

12) All Spain sites versus all northern Argentina sites
```{r echo = TRUE, eval = TRUE}

struc_tmp <- struc %>%
  inner_join(
    meta_dats %>%
      dplyr::select(Sample_Name, Population, Country, Site_Code),
    by = c("indiv" = "Sample_Name")
  ) %>%
  dplyr::mutate(
    pop = factor(Population, levels = c("Native", "Invasive")),
    pop = as.numeric(pop)
  ) %>%
  filter(Country == "Spain" | Country == "Argentina" & !Site_Code %in% c("ALGA", "BAIR", "LURO")) %>%
  dplyr::select(-c(Population, Country, Site_Code))

dim(struc_tmp) 
nrow(struc_tmp)/2 # 58 individuals

# Looks good
glimpse(struc_tmp[, 1:10])
struc_tmp %>%
  dplyr::select(indiv, pop) %>%
  View()

# Write a new file
file_nm <- "BS_12_SpainVersusNorthArgentina.str"
file.remove(file.path(file.path(res_path, "merged"), file_nm))

header <- c("\t","\t", paste(paste(mrkrs, collapse = " "), "\n", sep = ""))

cat(header, file = file.path(file.path(res_path, "merged"), file_nm), append = TRUE)

invisible(pbsapply(1:nrow(struc_tmp), function(i){
  
  cat(c(struc_tmp$indiv[i], "\t", struc_tmp$pop[i], "\t", paste(paste(struc_tmp[i, -grep("^indiv$|^pop$", names(struc_tmp))], collapse = " "), "\n", sep = "")), file = file.path(file.path(res_path, "merged"), file_nm), append = TRUE)
  
}))

# Opened the file in Vim to doublecheck structure, looks good

```

13) All Spain sites versus all southern Argentina sites
```{r echo = TRUE, eval = TRUE}

struc_tmp <- struc %>%
  inner_join(
    meta_dats %>%
      dplyr::select(Sample_Name, Population, Country, Site_Code),
    by = c("indiv" = "Sample_Name")
  ) %>%
  dplyr::mutate(
    pop = factor(Population, levels = c("Native", "Invasive")),
    pop = as.numeric(pop)
  ) %>%
  filter(Country == "Spain" | Country == "Argentina" & Site_Code %in% c("ALGA", "BAIR", "LURO")) %>%
  dplyr::select(-c(Population, Country, Site_Code))

dim(struc_tmp) 
nrow(struc_tmp)/2 # 66 individuals

# Looks good
glimpse(struc_tmp[, 1:10])
struc_tmp %>%
  dplyr::select(indiv, pop) %>%
  View()

# Write a new file
file_nm <- "BS_13_SpainVersusSouthArgentina.str"
file.remove(file.path(file.path(res_path, "merged"), file_nm))

header <- c("\t","\t", paste(paste(mrkrs, collapse = " "), "\n", sep = ""))

cat(header, file = file.path(file.path(res_path, "merged"), file_nm), append = TRUE)

invisible(pbsapply(1:nrow(struc_tmp), function(i){
  
  cat(c(struc_tmp$indiv[i], "\t", struc_tmp$pop[i], "\t", paste(paste(struc_tmp[i, -grep("^indiv$|^pop$", names(struc_tmp))], collapse = " "), "\n", sep = "")), file = file.path(file.path(res_path, "merged"), file_nm), append = TRUE)
  
}))

# Opened the file in Vim to doublecheck structure, looks good

```

14) All United States versus all Spain sites
```{r echo = TRUE, eval = TRUE}

struc_tmp <- struc %>%
  inner_join(
    meta_dats %>%
      dplyr::select(Sample_Name, Country),
    by = c("indiv" = "Sample_Name")
  ) %>%
  filter(Country %in% c("United States", "Spain")) %>%
  dplyr::mutate(
    pop = factor(Country, levels = c("United States", "Spain")),
    pop = as.numeric(pop)
  ) %>%
  dplyr::select(-c(Country))

dim(struc_tmp) 
nrow(struc_tmp)/2 # 127 individuals

# Looks good
glimpse(struc_tmp[, 1:10])
struc_tmp %>%
  dplyr::select(indiv, pop) %>%
  View()

# Write a new file
file_nm <- "BS_14_UnitedStatesVersusSpain.str"
file.remove(file.path(file.path(res_path, "merged"), file_nm))

header <- c("\t","\t", paste(paste(mrkrs, collapse = " "), "\n", sep = ""))

cat(header, file = file.path(file.path(res_path, "merged"), file_nm), append = TRUE)

invisible(pbsapply(1:nrow(struc_tmp), function(i){
  
  cat(c(struc_tmp$indiv[i], "\t", struc_tmp$pop[i], "\t", paste(paste(struc_tmp[i, -grep("^indiv$|^pop$", names(struc_tmp))], collapse = " "), "\n", sep = "")), file = file.path(file.path(res_path, "merged"), file_nm), append = TRUE)
  
}))

# Opened the file in Vim to doublecheck structure, looks good

```

15) Each invasive site versus all native sites (9 comparisons). Exclude SEVI because this was just one individual.
```{r echo = TRUE, eval = TRUE}

inv_sites <- meta_dats %>% 
  filter(Population == "Invasive") %>%
  filter(Site_Code != "SEVI") %>%
  pull(Site_Code) %>%
  unique() %>%
  as.character()

inv_sites
length(inv_sites)

# Iterate over invasive range sites to create one file each
# i <- 1
invisible(pblapply(1:length(inv_sites), function(i){
  
  inv_struc_tmp <- struc %>% 
    inner_join(
      meta_dats %>%
        dplyr::select(Sample_Name, Site_Code),
      by = c("indiv" = "Sample_Name")
    ) %>%
    filter(Site_Code == inv_sites[i]) %>%
    dplyr::mutate(
      pop = 2
    ) %>%
    dplyr::select(-c(Site_Code))
  
  # glimpse(inv_struc_tmp[, 1:10])
  # inv_struc_tmp[, 1:10] %>%
    # View()
  
  nat_struc_tmp <- struc %>%
    inner_join(
      meta_dats %>%
        dplyr::select(Sample_Name, Population),
      by = c("indiv" = "Sample_Name")
    ) %>%
    filter(Population == "Native") %>%
    dplyr::mutate(
      pop = 1
    ) %>%
    dplyr::select(-c(Population))
  
   # glimpse(nat_struc_tmp[, 1:10])
   # nat_struc_tmp[, 1:10] %>%
    # View()
   
   # Combine native and invasive range data
   struc_tmp <- inv_struc_tmp %>%
     bind_rows(nat_struc_tmp)

  # dim(struc_tmp) 
  # nrow(struc_tmp)/2
  # struc_tmp[, 1:5] %>%
  #   View()
   
  # Write a new file
  file_nm <- paste("BS_15-", i, "_", inv_sites[i], "_", "VersusAllNative.str", sep = "")
  # file_nm
  file.remove(file.path(file.path(res_path, "merged"), file_nm))

  header <- c("\t","\t", paste(paste(mrkrs, collapse = " "), "\n", sep = ""))

  cat(header, file = file.path(file.path(res_path, "merged"), file_nm), append = TRUE)

  invisible(pbsapply(1:nrow(struc_tmp), function(i){
  
    cat(c(struc_tmp$indiv[i], "\t", struc_tmp$pop[i], "\t", paste(paste(struc_tmp[i, -grep("^indiv$|^pop$", names(struc_tmp))], collapse = " "), "\n", sep = "")), file = file.path(file.path(res_path, "merged"), file_nm), append = TRUE)
  
  }))
  
}))

# Opened each file in Vim to doublecheck structure, these look good

```

16) Each invasive site versus all Uruguay sites
```{r echo = TRUE, eval = TRUE}

inv_sites <- meta_dats %>% 
  filter(Population == "Invasive") %>%
  filter(Site_Code != "SEVI") %>%
  pull(Site_Code) %>%
  unique() %>%
  as.character()

inv_sites
length(inv_sites)

# Iterate over invasive range sites to create one file each
# i <- 1
invisible(pblapply(1:length(inv_sites), function(i){
  
  inv_struc_tmp <- struc %>% 
    inner_join(
      meta_dats %>%
        dplyr::select(Sample_Name, Site_Code),
      by = c("indiv" = "Sample_Name")
    ) %>%
    filter(Site_Code == inv_sites[i]) %>%
    dplyr::mutate(
      pop = 2
    ) %>%
    dplyr::select(-c(Site_Code))
  
  # glimpse(inv_struc_tmp[, 1:10])
  # inv_struc_tmp[, 1:10] %>%
    # View()
  
  nat_struc_tmp <- struc %>%
    inner_join(
      meta_dats %>%
        dplyr::select(Sample_Name, Country),
      by = c("indiv" = "Sample_Name")
    ) %>%
    filter(Country == "Uruguay") %>%
    dplyr::mutate(
      pop = 1
    ) %>%
    dplyr::select(-c(Country))
  
   # glimpse(nat_struc_tmp[, 1:10])
   # nat_struc_tmp[, 1:10] %>%
    # View()
   
   # Combine native and invasive range data
   struc_tmp <- inv_struc_tmp %>%
     bind_rows(nat_struc_tmp)

  # dim(struc_tmp) 
  # nrow(struc_tmp)/2
  # struc_tmp[, 1:5] %>%
  #   View()
   
  # Write a new file
  file_nm <- paste("BS_16-", i, "_", inv_sites[i], "_", "VersusUruguay.str", sep = "")
  # file_nm
  file.remove(file.path(file.path(res_path, "merged"), file_nm))

  header <- c("\t","\t", paste(paste(mrkrs, collapse = " "), "\n", sep = ""))

  cat(header, file = file.path(file.path(res_path, "merged"), file_nm), append = TRUE)

  invisible(pbsapply(1:nrow(struc_tmp), function(i){
  
    cat(c(struc_tmp$indiv[i], "\t", struc_tmp$pop[i], "\t", paste(paste(struc_tmp[i, -grep("^indiv$|^pop$", names(struc_tmp))], collapse = " "), "\n", sep = "")), file = file.path(file.path(res_path, "merged"), file_nm), append = TRUE)
  
  }))
  
}))

# Opened each file in Vim to doublecheck structure, these look good

```

17) Each invasive site versus all Argentina sites
```{r echo = TRUE, eval = TRUE}

inv_sites <- meta_dats %>% 
  filter(Population == "Invasive") %>%
  filter(Site_Code != "SEVI") %>%
  pull(Site_Code) %>%
  unique() %>%
  as.character()

inv_sites
length(inv_sites)

# Iterate over invasive range sites to create one file each
# i <- 1
invisible(pblapply(1:length(inv_sites), function(i){
  
  inv_struc_tmp <- struc %>% 
    inner_join(
      meta_dats %>%
        dplyr::select(Sample_Name, Site_Code),
      by = c("indiv" = "Sample_Name")
    ) %>%
    filter(Site_Code == inv_sites[i]) %>%
    dplyr::mutate(
      pop = 2
    ) %>%
    dplyr::select(-c(Site_Code))
  
  # glimpse(inv_struc_tmp[, 1:10])
  # inv_struc_tmp[, 1:10] %>%
    # View()
  
  nat_struc_tmp <- struc %>%
    inner_join(
      meta_dats %>%
        dplyr::select(Sample_Name, Country),
      by = c("indiv" = "Sample_Name")
    ) %>%
    filter(Country == "Argentina") %>%
    dplyr::mutate(
      pop = 1
    ) %>%
    dplyr::select(-c(Country))
  
   # glimpse(nat_struc_tmp[, 1:10])
   # nat_struc_tmp[, 1:10] %>%
    # View()
   
   # Combine native and invasive range data
   struc_tmp <- inv_struc_tmp %>%
     bind_rows(nat_struc_tmp)

  # dim(struc_tmp) 
  # nrow(struc_tmp)/2
  # struc_tmp[, 1:5] %>%
  #   View()
   
  # Write a new file
  file_nm <- paste("BS_17-", i, "_", inv_sites[i], "_", "VersusArgentina.str", sep = "")
  # file_nm
  file.remove(file.path(file.path(res_path, "merged"), file_nm))

  header <- c("\t","\t", paste(paste(mrkrs, collapse = " "), "\n", sep = ""))

  cat(header, file = file.path(file.path(res_path, "merged"), file_nm), append = TRUE)

  invisible(pbsapply(1:nrow(struc_tmp), function(i){
  
    cat(c(struc_tmp$indiv[i], "\t", struc_tmp$pop[i], "\t", paste(paste(struc_tmp[i, -grep("^indiv$|^pop$", names(struc_tmp))], collapse = " "), "\n", sep = "")), file = file.path(file.path(res_path, "merged"), file_nm), append = TRUE)
  
  }))
  
}))

# Opened each file in Vim to doublecheck structure, these look good

```

18) Each invasive site versus all northern Argentina sites
```{r echo = TRUE, eval = TRUE}

inv_sites <- meta_dats %>% 
  filter(Population == "Invasive") %>%
  filter(Site_Code != "SEVI") %>%
  pull(Site_Code) %>%
  unique() %>%
  as.character()

inv_sites
length(inv_sites)

# Iterate over invasive range sites to create one file each
# i <- 1
invisible(pblapply(1:length(inv_sites), function(i){
  
  inv_struc_tmp <- struc %>% 
    inner_join(
      meta_dats %>%
        dplyr::select(Sample_Name, Site_Code),
      by = c("indiv" = "Sample_Name")
    ) %>%
    filter(Site_Code == inv_sites[i]) %>%
    dplyr::mutate(
      pop = 2
    ) %>%
    dplyr::select(-c(Site_Code))
  
  # glimpse(inv_struc_tmp[, 1:10])
  # inv_struc_tmp[, 1:10] %>%
    # View()
  
   nat_struc_tmp <- struc %>% 
     inner_join(
      meta_dats %>%
        dplyr::select(Sample_Name, Country, Site_Code),
        by = c("indiv" = "Sample_Name")
      ) %>%
      dplyr::mutate(
        pop = 1
      ) %>%
      filter(Country == "Argentina" & !Site_Code %in% c("ALGA", "BAIR", "LURO")) %>%
      dplyr::select(-c(Country, Site_Code))
  
   # glimpse(nat_struc_tmp[, 1:10])
   # nat_struc_tmp[, 1:10] %>%
    # View()
   
   # Combine native and invasive range data
   struc_tmp <- inv_struc_tmp %>%
     bind_rows(nat_struc_tmp)

  # dim(struc_tmp) 
  # nrow(struc_tmp)/2
  # struc_tmp[, 1:5] %>%
  #   View()
   
  # Write a new file
  file_nm <- paste("BS_18-", i, "_", inv_sites[i], "_", "VersusNorthArgentina.str", sep = "")
  # file_nm
  file.remove(file.path(file.path(res_path, "merged"), file_nm))

  header <- c("\t","\t", paste(paste(mrkrs, collapse = " "), "\n", sep = ""))

  cat(header, file = file.path(file.path(res_path, "merged"), file_nm), append = TRUE)

  invisible(pbsapply(1:nrow(struc_tmp), function(i){
  
    cat(c(struc_tmp$indiv[i], "\t", struc_tmp$pop[i], "\t", paste(paste(struc_tmp[i, -grep("^indiv$|^pop$", names(struc_tmp))], collapse = " "), "\n", sep = "")), file = file.path(file.path(res_path, "merged"), file_nm), append = TRUE)
  
  }))
  
}))

# Opened each file in Vim to doublecheck structure, these look good

```

19) Each invasive site versus all southern Argentina sites
```{r echo = TRUE, eval = TRUE}

inv_sites <- meta_dats %>% 
  filter(Population == "Invasive") %>%
  filter(Site_Code != "SEVI") %>%
  pull(Site_Code) %>%
  unique() %>%
  as.character()

inv_sites
length(inv_sites)

# Iterate over invasive range sites to create one file each
# i <- 1
invisible(pblapply(1:length(inv_sites), function(i){
  
  inv_struc_tmp <- struc %>% 
    inner_join(
      meta_dats %>%
        dplyr::select(Sample_Name, Site_Code),
      by = c("indiv" = "Sample_Name")
    ) %>%
    filter(Site_Code == inv_sites[i]) %>%
    dplyr::mutate(
      pop = 2
    ) %>%
    dplyr::select(-c(Site_Code))
  
  # glimpse(inv_struc_tmp[, 1:10])
  # inv_struc_tmp[, 1:10] %>%
    # View()
  
   nat_struc_tmp <- struc %>% 
     inner_join(
      meta_dats %>%
        dplyr::select(Sample_Name, Country, Site_Code),
        by = c("indiv" = "Sample_Name")
      ) %>%
      dplyr::mutate(
        pop = 1
      ) %>%
      filter(Country == "Argentina" & Site_Code %in% c("ALGA", "BAIR", "LURO")) %>%
      dplyr::select(-c(Country, Site_Code))
  
   # glimpse(nat_struc_tmp[, 1:10])
   # nat_struc_tmp[, 1:10] %>%
    # View()
   
   # Combine native and invasive range data
   struc_tmp <- inv_struc_tmp %>%
     bind_rows(nat_struc_tmp)

  # dim(struc_tmp) 
  # nrow(struc_tmp)/2
  # struc_tmp[, 1:5] %>%
  #   View()
   
  # Write a new file
  file_nm <- paste("BS_19-", i, "_", inv_sites[i], "_", "VersusSouthArgentina.str", sep = "")
  # file_nm
  file.remove(file.path(file.path(res_path, "merged"), file_nm))

  header <- c("\t","\t", paste(paste(mrkrs, collapse = " "), "\n", sep = ""))

  cat(header, file = file.path(file.path(res_path, "merged"), file_nm), append = TRUE)

  invisible(pbsapply(1:nrow(struc_tmp), function(i){
  
    cat(c(struc_tmp$indiv[i], "\t", struc_tmp$pop[i], "\t", paste(paste(struc_tmp[i, -grep("^indiv$|^pop$", names(struc_tmp))], collapse = " "), "\n", sep = "")), file = file.path(file.path(res_path, "merged"), file_nm), append = TRUE)
  
  }))
  
}))

# Opened each file in Vim to doublecheck structure, these look good

```

20) Randomizations to assess whether outlier loci detected were just as likely to have been detected by chance. As in Lemay & Russello 2015, randomly place all individuals into one of two groups, do this five times.
```{r echo = TRUE, eval = TRUE}

iter <- 5

# Randomly select 5 seeds for the random sampling below, such that the randomization can be replicated
set.seed(seed)
seeds <- sample(100, 5, replace = FALSE)
seeds

# i <- 1
invisible(pblapply(1:iter, function(i){
  
  set.seed(seeds[i])
  
  indivs_rsamp <- struc %>%
    pull(indiv) %>%
    unique() %>%
    as.character() %>%
    sample(length(.)/2, replace = FALSE)
  
  struc_tmp <- struc %>% 
    dplyr::mutate(
      indiv = as.character(indiv),
      pop = ifelse(indiv %in% indivs_rsamp, 1, 2)
    )
  
  # glimpse(struc_tmp[, 1:5])
  # struc_tmp[, 1:5] %>%
    # View()

  # Write a new file
  file_nm <- paste("BS_20-", i, "_RandomContrast.str", sep = "")
  # file_nm
  
  file.remove(file.path(file.path(res_path, "merged"), file_nm))

  header <- c("\t","\t", paste(paste(mrkrs, collapse = " "), "\n", sep = ""))

  cat(header, file = file.path(file.path(res_path, "merged"), file_nm), append = TRUE)

  invisible(pbsapply(1:nrow(struc_tmp), function(i){
  
    cat(c(struc_tmp$indiv[i], "\t", struc_tmp$pop[i], "\t", paste(paste(struc_tmp[i, -grep("^indiv$|^pop$", names(struc_tmp))], collapse = " "), "\n", sep = "")), file = file.path(file.path(res_path, "merged"), file_nm), append = TRUE)
  
  }))
  
}))

# Opened the files in Vim to doublecheck structure, these look good

# 64 Bayescan files, including 5 random contrasts as validation
length(list.files(path = file.path(file.path(res_path, "merged")), pattern = "BS_*"))

```

Next steps: positive control validations, remove positive controls from the filtered Structure file used as input above, remake BayeScan files, do PGDSpider conversions, start BayeScan runs, start BLASTing to mitochondrial genome as well
