---
title: "Sampling map"
author: "Grace Smith-Vidaurre"
date: "October 18, 2020"
output: html_document
---

```{r setup, eval = TRUE, echo = FALSE}

knitr::opts_knit$set(root.dir = "/home/owner/Desktop/GitHub_repos/origins-selection")

```

Purpose: Make sampling maps for RAD-seq thesis chapter manuscript.

```{r echo = TRUE, eval = TRUE}

rm(list = ls())

# "shapefiles", "broom", "geosphere", "GISTools",
X <- c("ggmap", "scales", "RColorBrewer", "rgeos", "rgdal", "sp", "pbapply", "shapefiles", "maps", "mapdata", "maptools", "measurements", "data.table", "grid", "gridExtra", "gtable", "plotrix", "gplots", "ggsn", "png", "colorspace", "grDevices", "ggplot2", "tidyverse")

invisible(lapply(X, library, character.only = TRUE))

gpath <- "/home/owner/Desktop/MANUSCRIPTS/Origins_Selection/GRAPHICS"

xls_path <- "/home/owner/Desktop/MANUSCRIPTS/Origins_Selection/DATA/Metadata_Barcodes"

```

Read in metadata.
```{r echo = TRUE, eval = TRUE}

meta_dats <- read.xlsx(file.path(xls_path, "Mymon_RADlibraries_2015_2019_CombinedMetadata.xlsx"))
glimpse(meta_dats)

```

Which sampling sites do not have geographic coordinates?
```{r echo = TRUE, eval = FALSE}

meta_dats %>%
  filter(is.na(Latitude_DD)) %>%
  pull(Site_Code) %>%
  unique()

```

Which sampling sites do have geographic coordinates? Note that some Connecticut (CNCT) and all Entre Rios (ERIO) samples did have coordinates, thanks to data passed on by Mike Russello.
```{r echo = TRUE, eval = FALSE}

meta_dats %>%
  filter(!is.na(Latitude_DD)) %>%
  pull(Site_Code) %>%
  unique()

```

Need to obtain geographic coordinates for these sites. Pull these from Google using the ggmap package.  
```{r echo = TRUE, eval = FALSE}

?mutate_geocode

# Make a data frame for geocoding with the sampling site location and country
geo_df <- meta_dats %>%
  filter(is.na(Latitude_DD)) %>%
  dplyr::mutate(
    Loc = paste(Sampling_Site, Country, sep = ", ")  
  ) %>%
  dplyr::select(Loc) %>%
  distinct()

geo_df

# Have to register on Google (AP key) to get this...and might need to provide a credit card to do so.
?register_google

geo_coords_df <- mutate_geocode(geo_df, Loc)

```



```{r echo = TRUE, eval = FALSE}

# 2. Read in trapping data and geographic coordinates

# waypts <- read.csv(file.path(path, "all_waypoints_manualedit.csv"), header = TRUE)
# str(waypts)
# 
# unique(waypts$Name)

smpls <- read.csv(file.path(path, "ParakeetBandingData_MASTER_GSV_28Nov17.csv"), header = TRUE)
str(smpls)
unique(smpls$Site.ID)

# 3. Merge the correct site coordinates with trapping data

# Making a match data frame manually, this needs to be fixed later
coords_df <- data.frame(site = c("ELGE", "SOLE", "SEMI-3", "EMBR", "1145", "CHAC", "PEIX", "BCAR", "1135", "BAGU"), lat = c(-34.427667, -34.425267, -34.3280918, -34.4444814, -34.3770252, -34.412789, -34.764865, -34.788002, -34.352689, -34.848051), lon = c(-57.812129, -57.743541, -57.6802826, -57.7288038, -57.5021464, -57.843938, -56.279224, -56.223475, -57.402123, -56.383672), department = c(rep("Colonia", 6), "Montevideo", "Montevideo", "Colonia", "Montevideo"))

# make site codes consistent
smpls$Site.ID <- str_replace(smpls$Site.ID, pattern = "1-145", replacement = "1145")
unique(smpls$Site.ID)

# merge data frames
smpls <- smpls %>%
  left_join(
    coords_df,
    by = c("Site.ID" = "site")
  )
str(smpls)
# View(smpls)

# 4. Make a map of genetic sampling areas

# Uruguay county border
URY_adm0 <- readOGR(dsn = path, layer = "URY_adm0")
slotNames(URY_adm0)
proj4string(URY_adm0)

# Uruguay county borders
URY_adm1 <- readOGR(dsn = path, layer = "URY_adm1")
slotNames(URY_adm1)
proj4string(URY_adm1)

# merge the URY county polgyons into one
str(URY_adm1@data)
m <- raster::union(URY_adm1[URY_adm1$NAME_1 == "Colonia", ], URY_adm1[URY_adm1$NAME_1 == "San José", ])
m2 <- raster::union(m, URY_adm1[URY_adm1$NAME_1 == "Canelones", ])
m3 <- raster::union(m2, URY_adm1[URY_adm1$NAME_1 == "Montevideo", ])
m4 <- raster::union(m3, URY_adm1[URY_adm1$NAME_1 == "Maldonado", ])
m5 <- raster::union(m4, URY_adm1[URY_adm1$NAME_1 == "Rocha", ])
m6 <- raster::union(m5, URY_adm1[URY_adm1$NAME_1 == "Salto", ])

# Read in Argentina and Brazil country borders
a <- readRDS(file.path(path, "gadm36_ARG_0_sp.rds"))
b <- readRDS(file.path(path, "gadm36_BRA_0_sp.rds"))

# order departments to match the geographic transect
smpls$department <- factor(smpls$department, levels = c("Colonia", "San Jose", "Salto", "Canelones", "Montevideo", "Maldonado", "Rocha"))
smpls$region <- smpls$department

smpls$region <- gsub("Canelones|Montevideo", "Canelones & Montevideo", smpls$region)
smpls$region <- factor(smpls$region, levels = c("Colonia", "San Jose", "Salto", "Canelones & Montevideo", "Maldonado", "Rocha"))

n <- 12
pie(rep(1,n), col = heat.colors(n))

# make colors for water and land fill colors
wfl <- gray.colors(n, alpha = 0.5)[12]
lfl <- "white"

# make sure colors sync across all figures
# ggplot recognizes unicode values for new symbols
# use these to encode department by both shapes and colors for greater clarity
fill.cols <- c(topo.colors(n, alpha = 0.75)[2], topo.colors(n, alpha = 0.9)[4], terrain.colors(n, alpha = 0.9)[1], gray.colors(n, alpha = 0.9)[2], gray.colors(n, alpha = 0.9)[2], heat.colors(n, alpha = 0.9)[6], heat.colors(n, alpha = 0.9)[1])
# shps <- c("\u25CF", "\u25C6", "\u25BC", "\u25A0", "\u25C4", "\u25B2")
# shps <- c(-as.hexmode("25CF"), -as.hexmode("25C6"), -as.hexmode("25BC"), -as.hexmode("25A0"), -as.hexmode("25C4"), -as.hexmode("25B2"))
cols <- c(rep("black", 5), heat.colors(n, alpha = 0.9)[6], "black")

shps <- c(21, 23, 25, 22, -as.hexmode("25C4"), 24)
sizes <- c(12, 12, 12, 12, 17, 12)

# library(Hmisc)
# show.col()
# show.pch()

TestUnicode <- function(start="25a0", end="25ff", ...)
{
  nstart <- as.hexmode(start)
  nend <- as.hexmode(end)
  r <- nstart:nend
  s <- ceiling(sqrt(length(r)))
  par(pty="s")
  plot(c(-1,(s)), c(-1,(s)), type="n", xlab="", ylab="",
       xaxs="i", yaxs="i")
  grid(s+1, s+1, lty=1)
  for(i in seq(r)) {
    try(points(i%%s, i%/%s, pch=-1*r[i],...))
  }
}

TestUnicode()
dev.off()

bufx <- 0.6
bufy <- 0.3

# Use these coordinates to make the same zoom map as in recordings figures
wp <- read.csv(file.path(path, "recording_waypoints.csv"), header = TRUE)
str(wp)

jpeg(file.path(file.path(gpath, "Uruguay_genetic_sampling_base_map_noscale.jpeg")), width = 27, height = 30, units = "cm", res = 300)

gg_base <- ggplot() +
  geom_polygon(aes(x = long, y = lat, group = group), color = "black",
               fill = lfl, size = 1, data = URY_adm0) + 
  geom_polygon(aes(x = long, y = lat, group = group), color = "black",
               fill = lfl, size = 0.6, data = a) + 
  geom_polygon(aes(x = long, y = lat, group = group), color = "black",
               fill = lfl, size = 0.6, data = b) + 
  geom_polygon(aes(x = long, y = lat, group = group), color = "black",
               fill = lfl, size = 0.6, data = m6) + 
  geom_point(aes(x = lon, y = lat, color = department, shape = region, size = region, fill = department), data = smpls) +
  scale_colour_manual(values = cols) + 
  scale_fill_manual(values = fill.cols) +
  scale_shape_manual(values = shps) + 
  scale_size_manual(values = sizes) + guides(color = FALSE, shape = FALSE, size = FALSE, fill = FALSE) +
  xlab("") + ylab("") + 
  theme(panel.background = element_rect(fill = wfl), plot.background = element_rect(fill = lfl), plot.title = element_text(size = 25, hjust = 0.5),
        panel.grid.major = element_line(size = 1, colour = wfl), 
        panel.grid.minor = element_line(size = 0.75, colour = wfl), 
        axis.line = element_line(size = 1.5, colour = "black"), 
        axis.title = element_text(size = 30), 
        axis.text = element_text(size = 30)) + 
  # coord_cartesian(xlim = c(min(smpls$lon) - bufx, max(smpls$lon) + bufx), y = c(min(smpls$lat) - bufy, max(smpls$lat) + bufy*3))
  coord_cartesian(xlim = c(min(wp$Longitude) - bufx, max(wp$Longitude) + bufx), y = c(min(wp$Latitude) - bufy, max(wp$Latitude) + bufy*3))

# gg_base

# gg_base <- gg_base + scale_bar(lon = -54.5, lat = -30.5,
#                                distance_lon = 50, distance_lat = 10, distance_legend = 50, 
#                                dist_unit = "km", orientation = FALSE, legend_size = 8)


buf <- 0.1
symbol <- 1
symbol <- sprintf("%02.f", symbol)
symbol <- readPNG(paste0(system.file("symbols", package = "ggsn"), 
                         "/", symbol, ".png"))
symbol <- rasterGrob(symbol, interpolate = TRUE)
gg_base <- gg_base + annotation_custom(grob = symbol, xmin = max(wp$Longitude) - 1.5, xmax = max(wp$Longitude) - buf, 
                                       ymin = max(wp$Latitude) - buf*4, ymax = max(wp$Latitude))

gg_base

dev.off()


# 5. Next, make a smaller inset map of the southwestern coast

buf1 <- 0.5
buf2 <- 0.05


jpeg(file.path(gpath, "Uruguay_sampling_inset_map_noscalebar.jpeg"), width = 25, height = 10, units = "cm", res = 300)

# [URY_adm1$NAME_1 == "Colonia"| URY_adm1$NAME_1 == "Montevideo", ]

gg_inset <- ggplot() +
  geom_polygon(aes(x = long, y = lat, group = group), color = "black",
               fill = "white", size = 0.5, data = URY_adm1[grepl("Colonia|San|Canelones|Montevideo|Maldonado|Rocha", URY_adm1$NAME_1)]) + 
  geom_point(aes(x = lon, y = lat, color = department, shape = region, size = region, fill = department), data = smpls, size = 6) +
  scale_colour_manual(values = cols) + 
  scale_fill_manual(values = fill.cols) +
  scale_shape_manual(values = shps) + 
  scale_size_manual(values = sizes) +
  xlab("") + ylab("") + 
  theme(panel.background = element_rect(fill = wfl), plot.background = element_rect(fill = "white"), plot.title = element_text(size = 25, hjust = 0.5),
        panel.grid.major = element_line(size = 1, colour = "white"), 
        panel.grid.minor = element_line(size = 0.75, colour = "white"), 
        axis.line = element_line(size = 0.75, colour = "black"), 
        axis.title = element_text(size = 32), 
        axis.text = element_text(size = 15)) + 
  
  coord_cartesian(xlim = c(min(wp$Longitude) - bufx, -55.5), y = c(-35.0, -34.2)) +
  
  # coord_cartesian(xlim = c(min(smpls$lon) - buf1, max(smpls$lon) + buf1), ylim = c(min(smpls$lat) - buf2, max(smpls$lat) + buf2)) +
  guides(color = FALSE, shape = FALSE)

gg_inset

dev.off()


# 6. Make a general worldwide RAD sampling map with state/country polygons filled in
n <- 12
states <- map_data("state") %>%
  filter(region %in% c("washington", "illinois", "florida", "connecticut"))
str(states)

spain <- map_data("world") %>%
  filter(region %in% c("Spain"))
str(spain)

arg <- map_data("world") %>%
  filter(region %in% c("Argentina"))
str(arg)

world <- map_data("world")

gg_wrld <- ggplot() +
  geom_polygon(data = world, aes(x = long, y = lat, group = group), fill = "gray65") +
  geom_polygon(data = states, aes(x = long, y = lat, group = group), fill = heat.colors(n)[5]) +
  geom_polygon(data = spain, aes(x = long, y = lat, group = group), fill = heat.colors(n)[5]) +
  geom_polygon(data = arg, aes(x = long, y = lat, group = group), fill = topo.colors(n)[3]) +
  xlab("Longitude (Decimal Degrees)") + ylab("Latitude (Decimal Degrees)") +
  theme_classic() +
  theme(axis.title = element_text(size = 18), axis.text = element_text(size = 14))

gg_wrld <- gg_wrld +
  scalebar(x.min = 0, x.max = 50, y.min = -50, y.max = -20, dist = 1500, dd2km = TRUE, model = "WGS84", height = 0.15, st.size = 3.5, st.dist = 0.15)

gg_wrld


buf <- 0.1
symbol <- 1
symbol <- sprintf("%02.f", symbol)
symbol <- readPNG(paste0(system.file("symbols", package = "ggsn"), 
                         "/", symbol, ".png"))
symbol <- rasterGrob(symbol, interpolate = TRUE)
gg_wrld <- gg_wrld + annotation_custom(grob = symbol, xmin = -180, xmax = -130, 
                                       ymin = -10, ymax = 10)

gg_wrld

ggsave(file.path(gpath, "RAD-seq_Round1_WorldMap.jpeg"), units = "cm", width = 26, height = 14, dpi = 200)  




```
