---
title: "Sampling map"
author: "Grace Smith-Vidaurre"
date: "October 18, 2020"
output: html_document
---

```{r setup, eval = TRUE, echo = FALSE}

knitr::opts_knit$set(root.dir = "/home/owner/Desktop/GitHub_repos/origins-selection")

```

Purpose: Make sampling maps for RAD-seq thesis chapter manuscript. Can I access eBird sightings for Jan 2018 to Oct 2020 to get all countries in which monks have been seen this year? In their invasive range.

```{r echo = TRUE, eval = TRUE}

# Fails with memory issue, aborted
# library(devtools)
# install_github("coolbutuseless/ggpattern")

rm(list = ls())

X <- c("ggmap", "scales", "RColorBrewer", "rgeos", "rgdal", "sp", "pbapply", "shapefiles", "maps", "mapdata", "maptools", "measurements", "data.table", "grid", "gridExtra", "gtable", "plotrix", "gplots", "ggsn", "png", "colorspace", "grDevices", "ggplot2", "tidyverse", "openxlsx", "auk")

invisible(lapply(X, library, character.only = TRUE))

# Path to URY polygons
path <- "/home/owner/Desktop/MANUSCRIPTS/Origins_Selection/DATA"

# Path to eBird data
epath <- "/media/owner/MYIOPSITTA/R/Origins_Selection"

gpath <- "/home/owner/Desktop/MANUSCRIPTS/Origins_Selection/FIGURES"

xls_path <- "/home/owner/Desktop/MANUSCRIPTS/Origins_Selection/DATA/Metadata_Barcodes"

```

Read in metadata.
```{r echo = TRUE, eval = TRUE}

meta_dats <- read.xlsx(file.path(xls_path, "Mymon_RADlibraries_2015_2019_CombinedMetadata.xlsx"))
glimpse(meta_dats)

```

Which sampling sites do not have geographic coordinates? Just BAIR after manually entering coordinates for sites that didn't have these, BAIR birds were sampled in Buenos Aires but obtained from further south.
```{r echo = TRUE, eval = FALSE}

meta_dats %>%
  filter(is.na(Latitude_DD)) %>%
  pull(Site_Code) %>%
  unique()

```

Which sampling sites do have geographic coordinates? Note that some Connecticut (CNCT) and all Entre Rios (ERIO) samples did have coordinates prior to my own geocoding, thanks to data passed on by Mike Russello.
```{r echo = TRUE, eval = FALSE}

meta_dats %>%
  filter(!is.na(Latitude_DD)) %>%
  pull(Site_Code) %>%
  unique()

```

eBird basic dataset request for Myiopsitta monachus in all regions, on 28 October 2020, for Jan. 2018 to October 2020.
```{r echo = TRUE, eval = FALSE}

# grep("Hawaii|Alaska", unique(countries$state))

countries <- read_ebd(file.path(epath, "ebd_monpar_201801_202010_relSep-2020.txt")) %>%
  pull(country) %>%
  unique()

countries

# Exclude countries in the native range
countries <- countries[-grep("Brazil|Argentina|Bolivia|Uruguay|Paraguay", countries)]
countries

# Replace "United States" with "USA", replace "United Kingdom" with UK
countries <- gsub("United States", "USA", countries)
countries <- gsub("United Kingdom", "UK", countries)
countries

```

Make a general worldwide RAD sampling map. Consider coloring country boundaries for emphasis.
```{r echo = TRUE, eval = FALSE}

# Add info about countries where monk parakeet sightings documented on eBird from Jan 2018 - Oct 2020
inv_countries <- map_data("world") %>%
  filter(region %in% countries) %>%
  # Remove islands or regions known to not have sightings (checked eBird map online for this date range)
  filter(!subregion %in% c("Hawaii", "Alaska", "Guam", "Channel Islands", "Northern Ireland", "Shetland Islands", "Scotland", "Wales")) %>%
  filter(!grepl("Island|Isla|Isola|Isle", subregion)) %>%
  droplevels()
glimpse(inv_countries)

inv_countries %>%
  filter(region == "UK") %>%
  pull(subregion) %>%
  unique()

# us <- map_data("world") %>%
#   filter(region %in% c("USA")) %>%
#   filter(!subregion %in% c("Hawaii", "Alaska", "Guam", "Puerto Rico"))
# str(us)

# states <- map_data("state") %>%
#   filter(region %in% c("washington", "illinois", "florida", "connecticut"))
# str(states)

spain <- map_data("world") %>%
  filter(region %in% c("Spain"))
str(spain)

# Needs to be added 
canaria <- map_data("world") %>%
  filter(region %in% c("Canary Islands")) %>%
  filter(subregion == "Gran Canaria")
str(canaria)

# Instead, read in the BirdLife species distribution map
mymon_distr <- readOGR(dsn = path, layer = "Myiopsitta_monachus")
# str(mymon_distr)
class(mymon_distr)
str(mymon_distr@data)
length(mymon_distr)
length(mymon_distr[[1]])

# Citation
# mymon_distr@data$CITATION

# Keep only the largest polygon, which corresponds to South America
wh <- which(mymon_distr@data$Shape_Area == max(mymon_distr@data$Shape_Area))

mymon_distr <- mymon_distr[wh, ]
length(mymon_distr)
str(mymon_distr)

world <- map_data("world")
glimpse(world)

nat_bord <- c("navy")
inv_bord <- c("darkgoldenrod1")

gg_wrld <- ggplot(data = world) +
  geom_polygon(data = world, aes(x = long, y = lat, group = group), fill = "gray75") +
  # Add the countries with eBird sightings
   geom_polygon(data = inv_countries, aes(x = long, y = lat, group = group), color = inv_bord, fill = alpha(inv_bord, 0.5)) +
  geom_polygon(data = canaria, aes(x = long, y = lat, group = group), color = inv_bord, fill = alpha(inv_bord, 0.5)) +
  geom_polygon(data = mymon_distr, aes(x = long, y = lat, group = group), color = nat_bord, fill = alpha(nat_bord, 0.5)) +
  xlab("Longitude (Decimal Degrees)") + ylab("Latitude (Decimal Degrees)") +
  coord_cartesian(xlim = c(-160, 176), ylim = c(-52, 80)) +
  theme_classic() +
  theme(
    axis.title = element_text(size = 14), 
    axis.text = element_text(size = 14),
    panel.border = element_rect(fill = alpha("white", 0), color = "black", size = 0.75)
 )

gg_wrld

gg_wrld <- gg_wrld +
  ggsn::scalebar(data = NULL, x.min = -160, x.max = -100, y.min = -35, y.max = -5, dist = 2000, transform = TRUE, dist_unit = "km", model = "WGS84", height = 0.10, st.size = 3.5, st.dist = 0.15, border.size = 0.15)

gg_wrld


buf <- 0.1
symbol <- 1
symbol <- sprintf("%02.f", symbol)
symbol <- readPNG(paste0(system.file("symbols", package = "ggsn"), 
                         "/", symbol, ".png"))
symbol <- rasterGrob(symbol, interpolate = TRUE)
gg_wrld <- gg_wrld + annotation_custom(grob = symbol, xmin = -150, xmax = -100, 
                                       ymin = -25, ymax = 0)

gg_wrld

ggsave(file.path(gpath, "OriginsSelection_Figure1_WorldMap.tiff"), units = "in", width = 11, height = 3.5, dpi = 300)

ggsave(file.path(gpath, "OriginsSelection_Figure1_WorldMap_lrg.tiff"), units = "in", width = 8.05, height = 5, dpi = 300)

dev.off()

```

Now need to make inset maps. Each map should have points, and labels for sampling sites above these.

Uruguay and Argentina. Add the species distribution instead of country polygons.
```{r echo = TRUE, eval = FALSE}

ury <- map_data("world") %>%
  filter(region %in% c("Uruguay"))
str(ury)

arg <- map_data("world") %>%
  filter(region %in% c("Argentina"))
str(arg)

tmp_pts <- meta_dats %>%
  filter(Country %in% c("Uruguay", "Argentina")) %>%
  filter(!is.na(Latitude_DD)) %>%
  dplyr::select(Site_Code, Latitude_DD, Longitude_DD) %>%
  distinct()
glimpse(tmp_pts)

ury_arg_inset <- ggplot(data = tmp_pts) +
  geom_polygon(data = world, aes(x = long, y = lat, group = group), fill = "gray85") +
  geom_polygon(data = ury, aes(x = long, y = lat, group = group), color = nat_bord, fill = "gray85", size = 0.25) +
  geom_polygon(data = arg, aes(x = long, y = lat, group = group), color = nat_bord, fill = "gray85", size = 0.25) +
  geom_point(data = tmp_pts, aes(x = Longitude_DD, y = Latitude_DD), fill = alpha(nat_bord, 1), shape = 21, size = 2) +
  xlab("") + ylab("") +
  coord_cartesian(xlim = c(-65, -53), ylim = c(-40, -24)) +
  theme_classic() +
  theme(
    axis.title = element_text(size = 12), 
    axis.text.y = element_text(size = 12),
    axis.text.x = element_text(size = 12, angle = 30, hjust = 1),
    panel.border = element_rect(fill = alpha("white", 0), color = "black", size = 0.75)
   )

ury_arg_inset

ury_arg_inset <- ury_arg_inset +
  ggsn::scalebar(data = NULL, x.min = -75, x.max = -54, y.min = -39.5, y.max = -35, dist = 200, transform = TRUE, dist_unit = "km", model = "WGS84", height = 0.15, st.size = 2.5, st.dist = 0.1, border.size = 0.15)

ury_arg_inset

ggsave(file.path(gpath, "OriginsSelection_Figure1_NativeInset.tiff"), units = "in", width = 3.60, height = 2.9, dpi = 300)

dev.off()

```

Just Uruguay, include exporter's aviaries....Using a finer-detailed country polygon with departmental polygons too.
```{r echo = TRUE, eval = FALSE}

# From the Google Map I made in My Maps in 2017:
# -34.8028801, -56.2582612
# Camino La Redención 2448

tmp_pts <- meta_dats %>%
  filter(Country %in% c("Uruguay")) %>%
  filter(!is.na(Latitude_DD)) %>%
  dplyr::select(Site_Code, Latitude_DD, Longitude_DD) %>%
  distinct() %>%
  bind_rows(
    data.frame(
      Site_Code = "Exporter's aviaries",
      Latitude_DD = -34.8028801, 
      Longitude_DD = -56.2582612
    )
  )
glimpse(tmp_pts)

# Uruguay county border
URY_adm0 <- readOGR(dsn = path, layer = "URY_adm0")
slotNames(URY_adm0)
proj4string(URY_adm0)

# Uruguay county borders
URY_adm1 <- readOGR(dsn = path, layer = "URY_adm1")
slotNames(URY_adm1)
proj4string(URY_adm1)

# merge the URY county polgyons into one
str(URY_adm1@data)
m <- raster::union(URY_adm1[URY_adm1$NAME_1 == "Colonia", ], URY_adm1[URY_adm1$NAME_1 == "San José", ])
m2 <- raster::union(m, URY_adm1[URY_adm1$NAME_1 == "Canelones", ])
m3 <- raster::union(m2, URY_adm1[URY_adm1$NAME_1 == "Montevideo", ])
m4 <- raster::union(m3, URY_adm1[URY_adm1$NAME_1 == "Florida", ])

# Read in Argentina country borders
a <- readRDS(file.path(path, "gadm36_ARG_0_sp.rds"))

ury_inset <- ggplot(data = tmp_pts) +
  # geom_polygon(data = world, aes(x = long, y = lat, group = group), fill = "gray85") +
  # geom_polygon(data = ury, aes(x = long, y = lat, group = group), color = nat_bord, fill = "gray85", size = 0.25) +
  geom_polygon(data = a, aes(x = long, y = lat, group = group), color = nat_bord, fill = "gray85", size = 0.25) +
  geom_polygon(data = m4, aes(x = long, y = lat, group = group), color = nat_bord, fill = "gray85", size = 0.25) +
  geom_point(data = tmp_pts %>%
                 filter(Site_Code != "Exporter's aviaries"), 
             aes(x = Longitude_DD, y = Latitude_DD), fill = alpha(nat_bord, 0.65), shape = 21, size = 2) +
  geom_point(data = tmp_pts %>%
                 filter(Site_Code == "Exporter's aviaries"), 
             aes(x = Longitude_DD, y = Latitude_DD), color = "firebrick", shape = 8, size = 2) +
  xlab("") + ylab("") +
  coord_cartesian(xlim = c(-57.8, -56.2), ylim = c(-34.92, -34.3)) +
  theme_classic() +
  theme(
    axis.title = element_text(size = 12), 
    axis.text.y = element_text(size = 12),
    axis.text.x = element_text(size = 12),
    panel.border = element_rect(fill = alpha("white", 0), color = "black", size = 0.75)
   )

ury_inset

ury_inset <- ury_inset +
  ggsn::scalebar(data = NULL, x.min = -57.75, x.max = -57.0, y.min = -34.85, y.max = -34.7, dist = 25, transform = TRUE, dist_unit = "km", model = "WGS84", height = 0.25, st.size = 3, st.dist = 0.25, border.size = 0.15)

ury_inset

ggsave(file.path(gpath, "OriginsSelection_Figure1_URYInset.tiff"), units = "in", width = 6.35, height = 2.85, dpi = 300)

dev.off()

```

U.S.
```{r echo = TRUE, eval = FALSE}

tmp_pts <- meta_dats %>%
  filter(Country %in% c("United States")) %>%
  filter(!is.na(Latitude_DD)) %>%
  dplyr::select(Site_Code, Latitude_DD, Longitude_DD) %>%
  distinct()
glimpse(tmp_pts)
tmp_pts

us_inset <- ggplot(data = tmp_pts) +
  geom_polygon(data = world, aes(x = long, y = lat, group = group), fill = "gray85") +
  geom_polygon(data = us, aes(x = long, y = lat, group = group), color = inv_bord, fill = "gray85", size = 0.5) +
  # State borders
  # geom_polygon(data = states, aes(x = long, y = lat, group = group), color = inv_bord, fill = "gray85", size = 0.25) +
  geom_point(data = tmp_pts, aes(x = Longitude_DD, y = Latitude_DD), color = "black", fill = alpha(inv_bord, 1), shape = 21, size = 2) +
  xlab("") + ylab("") +
  coord_cartesian(xlim = c(-123, -66), ylim = c(25, 49)) +
  theme_classic() +
  theme(
    axis.title = element_text(size = 12), 
    axis.text.y = element_text(size = 12),
    axis.text.x = element_text(size = 12),
    panel.border = element_rect(fill = alpha("white", 0), color = "black", size = 0.5)
   )

# us_inset

us_inset <- us_inset +
  ggsn::scalebar(data = NULL, x.min = -120, x.max = -105, y.min = 27, y.max = 35, dist = 800, transform = TRUE, dist_unit = "km", model = "WGS84", height = 0.15, st.size = 4, st.dist = 0.2, border.size = 0.15)

us_inset

ggsave(file.path(gpath, "OriginsSelection_Figure1_USInset.tiff"), units = "in", width = 5.25, height = 2.45, dpi = 300)

dev.off()

```

Spain.
```{r echo = TRUE, eval = FALSE}

tmp_pts <- meta_dats %>%
  filter(Country %in% c("Spain")) %>%
  filter(!is.na(Latitude_DD)) %>%
  dplyr::select(Site_Code, Latitude_DD, Longitude_DD) %>%
  distinct()
glimpse(tmp_pts)
tmp_pts

sp_inset <- ggplot(data = tmp_pts) +
  geom_polygon(data = world, aes(x = long, y = lat, group = group), fill = "gray85") +
  geom_polygon(data = spain, aes(x = long, y = lat, group = group), color = inv_bord, fill = "gray85", size = 0.5) +
  geom_polygon(data = canaria, aes(x = long, y = lat, group = group), color = inv_bord, fill = "gray85", size = 0.5) +
  geom_point(data = tmp_pts, aes(x = Longitude_DD, y = Latitude_DD), color = "black", fill = alpha(inv_bord, 1), shape = 21, size = 2) +
  xlab("") + ylab("") +
  coord_cartesian(xlim = c(-17, 5), ylim = c(28, 44)) +
  theme_classic() +
  theme(
    axis.title = element_text(size = 12), 
    axis.text.y = element_text(size = 12),
    axis.text.x = element_text(size = 12),
    panel.border = element_rect(fill = alpha("white", 0), color = "black", size = 0.5)
   )

# sp_inset

sp_inset <- sp_inset +
  ggsn::scalebar(data = NULL, x.min = -15, x.max = 3, y.min = 30, y.max = 35, dist = 300, transform = TRUE, dist_unit = "km", model = "WGS84", height = 0.15, st.size = 4, st.dist = 0.2, border.size = 0.15)

sp_inset

ggsave(file.path(gpath, "OriginsSelection_Figure1_SpainInset.tiff"), units = "in", width = 5.25, height = 2.45, dpi = 300)

dev.off()

```

