---
title: "Barcode file for Stacks demultiplexing"
author: "Grace Smith-Vidaurre"
date: "May 14, 2020"
output: html_document
---

```{r setup, eval = TRUE, echo = FALSE}

knitr::opts_knit$set(root.dir = "/home/owner/Desktop/GitHub_repos/origins-selection")

```

Purpose of this script? What came before?

Clean environment, load packages, set path.
```{r echo = TRUE, eval = TRUE, message = FALSE}

rm(list = ls())

X <- c("tidyverse")

invisible(lapply(X, library, character.only = TRUE))

path <- "/home/owner/Desktop/MANUSCRIPTS/Origins&Selection/DATA/Metadata_Barcodes"

```

Read in metadata.
```{r echo = TRUE, eval = TRUE}

meta_dats <- read.csv(file.path(path, "Mymon_RADlibraries_2015_2019_CombinedMetadata.csv"))
glimpse(meta_dats)

```

See Stacks manual for barcode file format: http://catchenlab.life.illinois.edu/stacks/manual/#clean

From the manual: "The barcode file is a very simple format — one barcode per line; if you want to rename the output files, the sample name prefix is provided in the second column."

Make a barcode file for Stacks::process_radtags per single-end and paired-end sets of libraries.

Single-end libraries.
```{r echo = TRUE, eval = TRUE}

tmp_df <- meta_dats %>%
  filter(Sequencing_Type == "Single-end") %>%
  droplevels()
glimpse(tmp_df)



```

Paired-end libraries.
```{r echo = TRUE, eval = TRUE}

tmp_df <- meta_dats %>%
  filter(Sequencing_Type == "Paired-end") %>%
  droplevels()
glimpse(tmp_df)


```