---
title: "Barcode file for Stacks demultiplexing"
author: "Grace Smith-Vidaurre"
date: "May 14, 2020"
output: html_document
---

```{r setup, eval = TRUE, echo = FALSE}

knitr::opts_knit$set(root.dir = "/home/owner/Desktop/GitHub_repos/origins-selection")

```

Purpose of this script? What came before?

Clean environment, load packages, set path.
```{r echo = TRUE, eval = TRUE, message = FALSE}

rm(list = ls())

X <- c("tidyverse", "pbapply")

invisible(lapply(X, library, character.only = TRUE))

# Switch to working on MYIOPSITTA, copied the .csv from this folder over
# path <- "/home/owner/Desktop/MANUSCRIPTS/Origins&Selection/DATA/Metadata_Barcodes"
path <- "/media/owner/MYIOPSITTA/R/Origins&Selection/barcodes"

```

Read in metadata.
```{r echo = TRUE, eval = TRUE}

meta_dats <- read.csv(file.path(path, "Mymon_RADlibraries_2015_2019_CombinedMetadata.csv"))
glimpse(meta_dats)

```

See Stacks manual for barcode file format: http://catchenlab.life.illinois.edu/stacks/manual/#clean

From the manual: "The barcode file is a very simple format — one barcode per line; if you want to rename the output files, the sample name prefix is provided in the second column."

Make a barcode file for Stacks::process_radtags per single-end and paired-end sets of libraries. These .txt files should be tab spaced.

Single-end libraries.
```{r echo = TRUE, eval = TRUE}

seq_type <- c("Single-end", "Paired-end")
seq_abb <- c("SE", "PE")

# x <- 1
invisible(pblapply(1:length(seq_type), function(i){
  
  tmp_df <- meta_dats %>%
    filter(Sequencing_Type == seq_type[i]) %>%
    droplevels()
  # glimpse(tmp_df)

  # Iterate over libraries
  libs <- tmp_df %>%
    pull(Library_Plate) %>%
    unique()
  # libs
  
  lapply(1:length(libs), function(x){
    
    tmp_df2 <- tmp_df %>%
      filter(Library_Plate == libs[x])
  
    # Initialize a new .txt file per library
    fileConn <- file(file.path(path, paste(seq_abb[i], "_barcodes_library", libs[x], ".txt", sep = "")))
  
    barcodes <- tmp_df2 %>%
      pull(Barcode_Sequence)
    
    sample_nms <- tmp_df2 %>%
      pull(Sample_Name)
    
    writeLines(paste(barcodes, sample_nms, sep = "\t"), fileConn)
    close(fileConn)
    
  })
  
}))


```