---
title: "Structure graphics"
author: "Grace Smith-Vidaurre"
date: "October 5, 2020"
output: html_document
---

```{r setup, eval = TRUE, echo = FALSE}

knitr::opts_knit$set(root.dir = "/home/owner/Desktop/GitHub_repos/origins-selection")

```

Purpose: Made graphics of population structure following Structure Bayesian analyses assigning individuals to distinct numbers of genetic clusters (1 - 26). pophelper version 2.3.0.

```{r echo = TRUE, eval = TRUE, message = FALSE}

rm(list = ls())

# library(devtools)
# install_github("royfrancis/pophelper")

X <- c("tidyverse", "adegenet", "pophelper", "openxlsx", "ggplot2", "gridExtra", "gtable", "pbapply")
invisible(lapply(X, library, character.only = TRUE))

# Path to the metadata spreadsheet
xls_path <- "/home/owner/Desktop/MANUSCRIPTS/Origins_Selection/DATA/Metadata_Barcodes"

# Path to Structure output, neutral merged SNPs
# path <- "/media/owner/MYIOPSITTA/R/Origins_Selection/Structure/StructureHavester_neutralSNPs_results"
path <- "/media/owner/MYIOPSITTA/R/Origins_Selection/Structure"

# Path to Stacks output, including the HWE filtered SNPs in Structure format 
res_path <- "/media/owner/MYIOPSITTA/R/Origins_Selection/stacks"

gpath <- "/home/owner/Desktop/MANUSCRIPTS/Origins_Selection/FIGURES"
seed <- 401

# Need to set working directory for some functions
setwd(path)

```

Read in metadata.
```{r echo = TRUE, eval = TRUE}

meta_dats <- read.xlsx(file.path(xls_path, "Mymon_RADlibraries_2015_2019_CombinedMetadata.xlsx"))
glimpse(meta_dats)

```

Read in Structure file for neutral SNPs.
```{r echo = TRUE, eval = TRUE}

file_nm <- "merged_HWE_missingData_filters_noPosControlDups_neutralSNPs.str"

neutral_snps <- read.structure(file.path(file.path(res_path, "merged"), file_nm), n.ind = 173, n.loc = 320, col.lab = 1, col.pop = 2, row.marknames = 1, onerowperind = FALSE, ask = FALSE, NA.char = "-9")
str(neutral_snps)

str(neutral_snps@tab)

# Populations are numbered consecutively by sampling site, but individuals are not ordered by populations
neutral_snps@pop

```

# Neutral SNPs (320)

Evanno method to evaluate "best" K.
```{r echo = TRUE, eval = TRUE}

# Choose runs
struc_files <- list.files(path = path, pattern = "merged_neutralSNPs_K", full.names = TRUE)
head(struc_files)

# Read in Structure results files
# One element per iteration of K in the resulting list
qlist <- readQ(struc_files, filetype = "structure")
str(qlist)
names(qlist)

# Tabulate runs, yields a data frame with one row per K and replicate
# For STRUCTURE runs, the tabulateQ() function produces additional columns namely number of loci, number of burn-ins, number of replicates (reps), estimated ln probability of data (elpd), mean value of ln likelihood (mvll) and variance of ln likelihood (vll). The table is sorted by loci, ind and K.
df1 <- tabulateQ(qlist)
glimpse(df1)

# Summarise runs, summarises over replicates of K, yields one row per K
df2 <- summariseQ(df1)
glimpse(df2)

# Find "best" K with Evanno method 
# elpdmean is mean estimated Ln probability...not NA for K1, so could be used to assess likelihood of K 1 versus K 2
sum_stats <- evannoMethodStructure(data = df2, exportplot = FALSE, returnplot = FALSE)
glimpse(sum_stats)

# Couldn't get the plot above to work, so using my own
sum_stats %>%
  # Drop k = 1 and k = 26, these naturally have NAs in the delta calculation
  filter(!is.na(deltaK)) %>%
  ggplot(aes(x = k, y = deltaK)) +
  geom_line() +
  geom_point(shape = 21, size = 3, fill = "grey65") +
  xlab("K") +
  ylab("Delta K") +
  scale_x_continuous(limits = c(2, 25), breaks = seq(2, 25, 1), labels = seq(2, 25, 1)) +
  scale_y_continuous(limits = c(0, 150), breaks = seq(0, 150, 25), labels = seq(0, 150, 25)) +
  theme_bw() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )

ggsave(file.path(gpath, "DeltaK_320neutralSNPs.jpeg"), units = "in", width = 5, height = 4)

# Now, estimated ln probability of data (elpd)
sum_stats %>%
  ggplot(aes(x = k, y = elpdmean)) +
  geom_line() +
  geom_point(shape = 21, size = 3, fill = "grey65") +
  xlab("K") +
  ylab("Mean estimated Ln probability") +
  scale_x_continuous(limits = c(1, 26), breaks = seq(1, 26, 1), labels = seq(1, 26, 1)) +
  theme_bw() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )

ggsave(file.path(gpath, "LnProb_320neutralSNPs.jpeg"), units = "in", width = 5, height = 4)

```

Make a barplot for the 10 Structure K2 runs.
```{r echo = TRUE, eval = TRUE}

# These are many of the individuals that had membership of to the first cluster in K2 inside the file K2.indfile. These are the Southern Argentina individuals, with some URY, NAR, and INV individuals having partial membership as well. Not sure why the CLUMPP graphics below don't reflect this.
dimnames(neutral_snps@tab)[[1]][c(8, 17, 18, 19, 75, 79, 83, 84, 88, 90, 126, 105, 118, 127, 132, 136, 160, 161, 164, 166, 167, 170, 173)]

```

Get metadata for sorting individuals in these barplots by region and sampling site.
```{r echo = TRUE, eval = TRUE}

path <- "/media/owner/MYIOPSITTA/R/Origins_Selection/Structure"

# Get sampling sites
sites <- sapply(1:nrow(neutral_snps@tab), function(i){
  meta_dats$Site_Code[grep(paste("^", dimnames(neutral_snps@tab)[[1]][i], "$", sep = ""), meta_dats$Sample_Name)]
})
head(sites)
unique(sites)

# Get sampling regions
regions <- sapply(1:nrow(neutral_snps@tab), function(i){
  meta_dats$Region[grep(paste("^", dimnames(neutral_snps@tab)[[1]][i], "$", sep = ""), meta_dats$Sample_Name)]
})
head(regions)

# Update labels in plot
regions <- gsub("South Central Uruguay", "South-Central\n Uruguay", regions)
regions <- gsub("ern ", "ern\n ", regions)

unique(regions)

all_regs <- c("Southern\n Argentina", "Northern\n Argentina", "South-Central\n Uruguay", "Southwestern\n Uruguay", "Spain", "Southern\n United States", "Northern\n United States")

fact_regions <- factor(regions, levels = all_regs)

# Same order as in Structure file (taken from BayeScan_post-processing.Rmd)
all_sites <- c("CHAC", "ELGE", "EMBR", "SEMI-3", "SOLE", "1135", "1145", "BAGU", "BCAR", "PEIX", "ALGA", "BAIR", "LURO", "BAZO", "ERIO", "BARC", "MADR", "MALL", "SEVI", "ZARA", "GRCA", "FLOR", "ILLI", "WASH", "CNCT")

fact_sites <- factor(sites, levels = all_sites)

```

Make barplots with CLUMPP results from Discovery run. From the pophelper vignette: "Individual labels are row names of the ‘qlist’ dataframes...The rownames of ‘qlist’ can also be modified by the user by adding custom individual labels." 

Loop over K = 2 to K = 10.
```{r echo = TRUE, eval = TRUE}

Ks <- seq(2, 10, 1)

# k <- 1
qdf_list <- invisible(pblapply(1:length(Ks), function(k){
  
  # From the pophelper CLUMPP runs on my local machine
  qlist_aligned <- readQ(file.path(file.path(path, "CLUMPP_results/clumpp_pophelper"), paste("pop_K", Ks[k], "-combined-merged.txt", sep = "")), filetype = "clumpp")
  # str(qlist_aligned)

  # Sorting individuals by region and site
  qlist_aligned[[1]]$regions <- as.numeric(fact_regions)
  qlist_aligned[[1]]$sites <- as.numeric(fact_sites)
  
  # Make individual labels as site-number for each individual
  qlist_aligned[[1]] <- qlist_aligned[[1]] %>%
    dplyr::mutate(
      char_sites = as.character(fact_sites)
    ) %>%
    group_by(char_sites) %>% 
    dplyr::mutate(count = row_number(char_sites)) %>%
    ungroup() %>%
    dplyr::mutate(
      count = str_pad(count, width = 2, pad = "0", side = "left"),
      row_ID = paste(char_sites, count, sep = "-")
    ) %>%
    as.data.frame()
  
  # Update rownames with the new column
  rownames(qlist_aligned[[1]]) <- qlist_aligned[[1]]$row_ID
  
  # Do the sorting
  ord <- order(qlist_aligned[[1]]$regions, qlist_aligned[[1]]$sites, decreasing = FALSE)
  qlist_aligned[[1]] <- qlist_aligned[[1]][ord, ]
  # head(qlist_aligned[[1]])
  
  # Get rid of the region and site columns after sorting
  qlist_aligned[[1]] <- qlist_aligned[[1]][, -grep("region|site|char_sites|count|row_ID", names(qlist_aligned[[1]]))]
  # head(qlist_aligned[[1]])

  # qlist_aligned[[1]]

  # Which samples have membership greater than 0.1 but less than 0.5 to Cluster 1?
  # Matches what I was seeing above, same CNCT individual for instance
  # qlist_aligned[[1]] %>%
    # filter(Cluster1 >= 0.1 & Cluster1 <= 0.5)

  # qlist_aligned[[1]] %>%
    # filter(Cluster1 >= 0.5)

  # From pophelper: One or more sets of group labels can be added to the bottom of barplots by providing a ‘data.frame’ object. The number of rows of the ‘data.frame’ must be equal to the number of individuals. The header of the ‘data.frame’ is the group label set title.

  # Convert back to qlist attributes
  qlist_aligned <- as.qlist(qlist_aligned)
  # str(qlist_aligned)
  
  return(qlist_aligned[[1]])
  
}))

str(qdf_list)

# Get qlist attributes back
qdf_list <- as.qlist(qdf_list)
str(qdf_list)
  
# Custom strip panel label showing k only
fn1 <- function(x) attr(x, "k")
spnames <- paste0("K = ", sapply(qdf_list, fn1))

# Region label, order this
lab_df <- data.frame(Region = regions) %>%
  dplyr::mutate(
    # Make region labels more general
    Region = gsub("Southern\n Argentina|Northern\n Argentina", "Argentina", Region),
    Region = gsub("South-Central\n Uruguay|Southwestern\n Uruguay", "Uruguay", Region),
    Region = gsub("Southern\n United States|Northern\n United States", "United States", Region),
    Region = factor(Region, levels = c("Argentina", "Uruguay", "Spain", "United States")),
    Site = factor(sites, levels = all_sites)
  ) %>%
  arrange(-desc(Region), -desc(Site)) %>%
  dplyr::mutate(
    Region = as.character(Region)
  ) %>%
  dplyr::select(-c(Site))
glimpse(lab_df)
lab_df

names(lab_df) <- " "

# When using all 10 K
# cols <- c("firebrick", "maroon", "royalblue", "gold2", "grey35", "coral", "slateblue1", "orangered", "black", "wheat")

# K of 2 to 5
cols <- c("firebrick", "royalblue", "gold2", "grey35", "coral")

# K of 6 - 10 don't provide much more information
# alignK aligns colors for clusters, at least for the first few K

# With individual labels
p <- plotQ(alignK(qdf_list[c(1:4)]), returnplot = TRUE, exportplot = FALSE, quiet = TRUE, basesize = 11, grplab = lab_df, grplabsize = 3, grplabcol = "black", grplabface = "bold", linesize = 27, pointsize = 6, linealpha = 0.2, pointcol = "white", grplabpos = 0.5, linepos = 0.5, divcol = "white", divsize = 0.55, grplabheight = 2.75, grplabangle = 0, panelratio = c(4, 2), clustercol = cols, splab = spnames[c(1:4)], spbgcol = "grey85", splabsize = 12, splabface = "bold", showindlab = TRUE, useindlab = TRUE, indlabangle = 90, indlabcol = "black", barbordersize = 0.1, barbordercolour = "white", indlabvjust = -0.5, indlabhjust = 0, indlabsize = 6, barsize = 1, theme = "theme_bw", imgoutput = "join", showyaxis = TRUE, panelspacer = 0.25, showlegend = TRUE, legendkeysize = 10, legendtextsize = 10, legendrow = 1, legendlab = as.character(seq(1, 5, 1)))
grid.arrange(p$plot[[1]])

rownames(qdf_list[1])

# With individual labels
p <- plotQ(alignK(qdf_list[c(1:4)]), returnplot = TRUE, exportplot = FALSE, quiet = TRUE, basesize = 12, grplab = lab_df, grplabsize = 6, grplabcol = "black", grplabface = "bold", linesize = 10, grplabspacer = 0, pointsize = 6, linealpha = 0.2, pointcol = "white", grplabpos = 0.5, linepos = 0.5, divcol = "white", divsize = 0.55, grplabheight = 2.75, grplabangle = 0, panelratio = c(4, 1), clustercol = cols, splab = spnames[c(1:4)], spbgcol = "grey85", splabcol = "black", splabsize = 12, splabface = "bold", showindlab = FALSE, useindlab = TRUE, indlabangle = 90, indlabcol = "black", barbordersize = 0.1, barbordercolour = "white", indlabvjust = -0.5, indlabhjust = 0, indlabsize = 6, barsize = 1, theme = "theme_bw", imgoutput = "join", showyaxis = TRUE, panelspacer = 0.25, showlegend = TRUE, legendkeysize = 4, legendtextsize = 12, legendrow = 1, legendlab = as.character(seq(1, 5, 1)))
dev.off()
  
# Write out image file, need to put a white box over the 6 in legend, the function fails when trying to use 2-5 for legend labels
tiff(file.path(gpath, "Structure_neutral_K2toK5.tiff"), units = "in", width = 8.5, height = 6, res = 600)
grid.arrange(p$plot[[1]])
dev.off()

# Write out image file
# Doesn't work as expected
# plotQ(alignK(qdf_list[c(1:4)]), returnplot = FALSE, exportplot = TRUE, quiet = TRUE, basesize = 11, grplab = lab_df, grplabsize = 4, grplabcol = "black", grplabface = "bold", linesize = 27, pointsize = 6, linealpha = 0.2, pointcol = "white", grplabpos = 0.5, linepos = 0.5, divcol = "white", divsize = 0.75, grplabheight = 2.75, grplabangle = 0, panelratio = c(4, 2), clustercol = cols, splab = spnames[c(1:4)], spbgcol = "grey85", splabsize = 12, splabface = "bold", showindlab = TRUE, useindlab = TRUE, indlabangle = 90, indlabcol = "black", barbordersize = 0.1, barbordercolour = "white", indlabvjust = 0.5, indlabhjust = 0, indlabsize = 8, barsize = 1, theme = "theme_bw", imgoutput = "join", showyaxis = TRUE, panelspacer = 0.25, showlegend = TRUE, legendkeysize = 10, legendtextsize = 10, legendrow = 1, legendlab = as.character(seq(1, 5, 1)), exportpath = gpath, imgtype = "tiff", outputfilename = "Structure_K2toK5.tiff", dpi = 600)


```

To do:

- Add this to LibreOffice and add sampling site labels by hand.


I tried running pophelper with the CLUMPP output from Discovery, but it didn't recognize the indq file for K2.
```{r echo = TRUE, eval = TRUE}

# From Discovery runs
# # Read aligned K2 population file
# aligned <- readQ(file.path(file.path(path, "CLUMPP_results"), "K2.popq"), filetype = "clumpp")
# str(aligned)
# 
# # Read aligned K2 individual file
# aligned <- readQ(file.path(file.path(path, "CLUMPP_results"), "K2.indq"), filetype = "clumpp")
# str(aligned)

```


# Outlier SNPs (241)

Read in Structure file for the 241 outlier SNPs. Here, ran only K 1 through 10.
```{r echo = TRUE, eval = TRUE}

file_nm <- "merged_HWE_missingData_filters_noPosControlDups_outlierSNPs.str"

outlier_snps <- read.structure(file.path(file.path(res_path, "merged"), file_nm), n.ind = 173, n.loc = 241, col.lab = 1, col.pop = 2, row.marknames = 1, onerowperind = FALSE, ask = FALSE, NA.char = "-9")
str(outlier_snps)

str(outlier_snps@tab)

# Populations are numbered consecutively by sampling site, but individuals are not ordered by populations
outlier_snps@pop

```

Evanno method to evaluate "best" K.
```{r echo = TRUE, eval = TRUE}

path <- "/media/owner/MYIOPSITTA/R/Origins_Selection/Structure_outliers"

# Choose runs
struc_files <- list.files(path = path, pattern = "merged_outlierSNPs_K", full.names = TRUE)
head(struc_files)

# Read in Structure results files
# One element per iteration of K in the resulting list
qlist <- readQ(struc_files, filetype = "structure")
str(qlist)
names(qlist)

# Tabulate runs, yields a data frame with one row per K and replicate
# For STRUCTURE runs, the tabulateQ() function produces additional columns namely number of loci, number of burn-ins, number of replicates (reps), estimated ln probability of data (elpd), mean value of ln likelihood (mvll) and variance of ln likelihood (vll). The table is sorted by loci, ind and K.
df1 <- tabulateQ(qlist)
glimpse(df1)

# Summarise runs, summarises over replicates of K, yields one row per K
df2 <- summariseQ(df1)
glimpse(df2)

# Find "best" K with Evanno method 
# elpdmean is mean estimated Ln probability...not NA for K1, so could be used to assess likelihood of K 1 versus K 2
sum_stats <- evannoMethodStructure(data = df2, exportplot = FALSE, returnplot = FALSE)
glimpse(sum_stats)

# Couldn't get the plot above to work, so using my own
sum_stats %>%
  # Drop k = 1 and 10, naturally have NAs in the delta calculation
  filter(!is.na(deltaK)) %>%
  ggplot(aes(x = k, y = deltaK)) +
  geom_line() +
  geom_point(shape = 21, size = 3, fill = "grey65") +
  xlab("K") +
  ylab("Delta K") +
  scale_x_continuous(limits = c(2, 9), breaks = seq(2, 9, 1), labels = seq(2, 9, 1)) +
  scale_y_continuous(limits = c(0, 600), breaks = seq(0, 600, 50), labels = seq(0, 600, 50)) +
  theme_bw() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )

ggsave(file.path(gpath, "DeltaK_241outlierSNPs.jpeg"), units = "in", width = 5, height = 4)


# Now, estimated ln probability of data (elpd)
sum_stats %>%
  ggplot(aes(x = k, y = elpdmean)) +
  geom_line() +
  geom_point(shape = 21, size = 3, fill = "grey65") +
  xlab("K") +
  ylab("Mean estimated Ln probability") +
  scale_x_continuous(limits = c(1, 10), breaks = seq(1, 10, 1), labels = seq(1, 10, 1)) +
  theme_bw() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )

ggsave(file.path(gpath, "LnProb_241outlierSNPs.jpeg"), units = "in", width = 5, height = 4)

```

Get metadata for sorting individuals in CLUMPP runs to make barplots by region and sampling site.
```{r echo = TRUE, eval = TRUE}

# Get sampling sites
sites <- sapply(1:nrow(outlier_snps@tab), function(i){
  meta_dats$Site_Code[grep(paste("^", dimnames(outlier_snps@tab)[[1]][i], "$", sep = ""), meta_dats$Sample_Name)]
})
head(sites)
unique(sites)

# Get sampling regions
regions <- sapply(1:nrow(outlier_snps@tab), function(i){
  meta_dats$Region[grep(paste("^", dimnames(outlier_snps@tab)[[1]][i], "$", sep = ""), meta_dats$Sample_Name)]
})
head(regions)

# Update labels in plot
regions <- gsub("South Central Uruguay", "South-Central\n Uruguay", regions)
regions <- gsub("ern ", "ern\n ", regions)

unique(regions)

all_regs <- c("Southern\n Argentina", "Northern\n Argentina", "South-Central\n Uruguay", "Southwestern\n Uruguay", "Spain", "Southern\n United States", "Northern\n United States")

fact_regions <- factor(regions, levels = all_regs)

# Same order as in Structure file (taken from BayeScan_post-processing.Rmd)
all_sites <- c("CHAC", "ELGE", "EMBR", "SEMI-3", "SOLE", "1135", "1145", "BAGU", "BCAR", "PEIX", "ALGA", "BAIR", "LURO", "BAZO", "ERIO", "BARC", "MADR", "MALL", "SEVI", "ZARA", "GRCA", "FLOR", "ILLI", "WASH", "CNCT")

fact_sites <- factor(sites, levels = all_sites)

```

Use pophelper to summarize across replicates to make CLUMPP input. Note that I did this for the neutral SNPs too but I think I foolishly deleted that code. Results in subdirectories inside the path called "pop_K*".
```{r echo = TRUE, eval = FALSE}

path <- "/media/owner/MYIOPSITTA/R/Origins_Selection/Structure_outliers"

# Set working directory again since the function below doesn't take a path
setwd(path)
# getwd()

# Read in Structure K replicates as a qlist
outlier_files <- list.files(path, pattern = "merged_outlierSNPs_K", full.names = TRUE)
str(outlier_files)

outlier_qlist <- readQ(outlier_files)
str(outlier_qlist)

# Use this function to export files to be used for a CLUMPP run on my local machine
# Used 2 for neutral runs as well, and 20 iterations
clumppExport(qlist = outlier_qlist, prefix = "pop", parammode = 2,
paramrep = 20)

```

After making these new directories with CLUMPP input files, I updated the script "run_CLUMPP" and ran this for outlier loci on my local machine. All the output files get saved to tje directory where the code is saved, despite setting the working directory, so I had to move these into the right folder for making graphics (see below).

Looped over K = 2 to K = 10 to read in CLUMPP results and make Structure raphics across iterations per K. 
```{r echo = TRUE, eval = TRUE}

# Made the nested folders CLUMPP_results/clumpp_pophelper inside the directory below

path <- "/media/owner/MYIOPSITTA/R/Origins_Selection/Structure_outliers"

Ks <- seq(2, 9, 1) # Read files up to K9 as K10 still running in CLUMPP

# k <- 1
qdf_list <- invisible(pblapply(1:length(Ks), function(k){
  
  # From the pophelper CLUMPP runs on my local machine
  qlist_aligned <- readQ(file.path(file.path(path, "CLUMPP_results/clumpp_pophelper"), paste("pop_K", Ks[k], "-combined-merged.txt", sep = "")), filetype = "clumpp")
  # str(qlist_aligned)

  # Sorting individuals by region and site
  qlist_aligned[[1]]$regions <- as.numeric(fact_regions)
  qlist_aligned[[1]]$sites <- as.numeric(fact_sites)
  
  # Make individual labels as site-number for each individual
  qlist_aligned[[1]] <- qlist_aligned[[1]] %>%
    dplyr::mutate(
      char_sites = as.character(fact_sites)
    ) %>%
    group_by(char_sites) %>% 
    dplyr::mutate(count = row_number(char_sites)) %>%
    ungroup() %>%
    dplyr::mutate(
      count = str_pad(count, width = 2, pad = "0", side = "left"),
      row_ID = paste(char_sites, count, sep = "-")
    ) %>%
    as.data.frame()
  
  # Update rownames with the new column
  rownames(qlist_aligned[[1]]) <- qlist_aligned[[1]]$row_ID
  
  # Do the sorting
  ord <- order(qlist_aligned[[1]]$regions, qlist_aligned[[1]]$sites, decreasing = FALSE)
  qlist_aligned[[1]] <- qlist_aligned[[1]][ord, ]
  # head(qlist_aligned[[1]])
  
  # Get rid of the region and site columns after sorting
  qlist_aligned[[1]] <- qlist_aligned[[1]][, -grep("region|site|char_sites|count|row_ID", names(qlist_aligned[[1]]))]
  # head(qlist_aligned[[1]])

  # qlist_aligned[[1]]

  # Which samples have membership greater than 0.1 but less than 0.5 to Cluster 1?
  # Matches what I was seeing above, same CNCT individual for instance
  # qlist_aligned[[1]] %>%
    # filter(Cluster1 >= 0.1 & Cluster1 <= 0.5)

  # qlist_aligned[[1]] %>%
    # filter(Cluster1 >= 0.5)

  # From pophelper: One or more sets of group labels can be added to the bottom of barplots by providing a ‘data.frame’ object. The number of rows of the ‘data.frame’ must be equal to the number of individuals. The header of the ‘data.frame’ is the group label set title.

  # Convert back to qlist attributes
  qlist_aligned <- as.qlist(qlist_aligned)
  # str(qlist_aligned)
  
  return(qlist_aligned[[1]])
  
}))

str(qdf_list)

# Get qlist attributes back
qdf_list <- as.qlist(qdf_list)
str(qdf_list)
  
# Custom strip panel label showing k only
fn1 <- function(x) attr(x, "k")
spnames <- paste0("K = ", sapply(qdf_list, fn1))

# Region label, order this
lab_df <- data.frame(Region = regions) %>%
  dplyr::mutate(
    # Make region labels more general
    Region = gsub("Southern\n Argentina|Northern\n Argentina", "Argentina", Region),
    Region = gsub("South-Central\n Uruguay|Southwestern\n Uruguay", "Uruguay", Region),
    Region = gsub("Southern\n United States|Northern\n United States", "United States", Region),
    Region = factor(Region, levels = c("Argentina", "Uruguay", "Spain", "United States")),
    Site = factor(sites, levels = all_sites)
  ) %>%
  arrange(-desc(Region), -desc(Site)) %>%
  dplyr::mutate(
    Region = as.character(Region)
  ) %>%
  dplyr::select(-c(Site))
glimpse(lab_df)
lab_df

names(lab_df) <- " "

length(Ks)

# When using all 10 K
# cols <- c("firebrick", "royalblue", "gold2", "grey35", "wheat", "coral", "slateblue1", "orangered", "black", "maroon")

# K of 2 to 5
cols <- c("royalblue", "firebrick", "gold2", "grey35", "coral")

# K of 6 and beyond don't provide much more information
# alignK aligns colors for clusters, at least for the first few K

# With individual labels, K 2 - 9, no legend
# p <- plotQ(alignK(qdf_list[c(1:8)]), returnplot = TRUE, exportplot = FALSE, quiet = TRUE, basesize = 12, grplab = lab_df, grplabsize = 6, grplabcol = "black", grplabface = "bold", linesize = 10, grplabspacer = 0, pointsize = 6, linealpha = 0.2, pointcol = "white", grplabpos = 0.5, linepos = 0.5, divcol = "white", divsize = 0.55, grplabheight = 2.75, grplabangle = 0, panelratio = c(4, 1), clustercol = cols, splab = spnames[c(1:8)], spbgcol = "grey85", splabcol = "black", splabsize = 12, splabface = "bold", showindlab = FALSE, useindlab = TRUE, indlabangle = 90, indlabcol = "black", barbordersize = 0.1, barbordercolour = "white", indlabvjust = -0.5, indlabhjust = 0, indlabsize = 6, barsize = 1, theme = "theme_bw", imgoutput = "join", showyaxis = TRUE, panelspacer = 0.25, showlegend = FALSE)
# grid.arrange(p$plot[[1]])
# 
# rownames(qdf_list[1])

# With individual labels
p <- plotQ(alignK(qdf_list[c(1:4)]), returnplot = TRUE, exportplot = FALSE, quiet = TRUE, basesize = 12, grplab = lab_df, grplabsize = 6, grplabcol = "black", grplabface = "bold", linesize = 10, grplabspacer = 0, pointsize = 6, linealpha = 0.2, pointcol = "white", grplabpos = 0.5, linepos = 0.5, divcol = "white", divsize = 0.55, grplabheight = 2.75, grplabangle = 0, panelratio = c(4, 1), clustercol = cols, splab = spnames[c(1:4)], spbgcol = "grey85", splabcol = "black", splabsize = 12, splabface = "bold", showindlab = FALSE, useindlab = TRUE, indlabangle = 90, indlabcol = "black", barbordersize = 0.1, barbordercolour = "white", indlabvjust = -0.5, indlabhjust = 0, indlabsize = 6, barsize = 1, theme = "theme_bw", imgoutput = "join", showyaxis = TRUE, panelspacer = 0.25, showlegend = TRUE, legendkeysize = 4, legendtextsize = 12, legendrow = 1, legendlab = as.character(seq(1, 5, 1)))
dev.off()
  
# Write out image file
tiff(file.path(gpath, "Structure_outliers_K2toK5.tiff"), units = "in", width = 8.5, height = 6, res = 600)
grid.arrange(p$plot[[1]])
dev.off()

```

Some interesting patterns here that are different to the neutral loci. K of 6 and beyond don't provide new information. 